\input{preamble}
\title{Distilled Mathematical Arguments from Alquier's PAC-Bayes Tutorial}
\author{Generated proof-target note}
\date{Source PDF: \texttt{alquier-2024\_user-friendly-introduction-to-pac-bayes\_\_arXiv-2110.11216.pdf}}
\begin{document}
\maketitle

\section*{Reference being distilled}
This note distills Alquier, \emph{User-Friendly Introduction to PAC-Bayes Bounds}, Foundations and Trends in Machine Learning 17(2), 174--303, 2024 \cite{Alquier2024}.

\section*{Role in the formalization}
This is a background source for KL-regularized variational bounds.  It is not a primary Schr\"odinger bridge or Sinkhorn convergence source, but its PAC-Bayes change-of-measure argument is mathematically adjacent to Donsker--Varadhan/Gibbs variational lemmas.

\begin{distillation}
\item \textbf{Change of measure.}  For a prior $\pi$ over hypotheses and posterior $\rho\ll\pi$, expectations under $\rho$ are controlled by a KL penalty $\KL(\rho\mid\pi)$ and an exponential moment under $\pi$.
\item \textbf{Gibbs variational inequality.}  A core reusable lemma is
\[
  \E_{h\sim\rho} f(h)
  \le \KL(\rho\mid\pi)+\log \E_{h\sim\pi}e^{f(h)}.
\]
This is the one-sided form of the Donsker--Varadhan variational formula.
\item \textbf{PAC conversion.}  Exponential concentration for each hypothesis is lifted to a high-probability statement uniform over all posteriors by integrating over the prior and applying the change-of-measure inequality.
\item \textbf{Optimization interpretation.}  The best posterior often has Gibbs form $d\rho^*/d\pi\propto e^{-\lambda L(h)}$, mirroring the entropic optimality conditions that appear in Sinkhorn/DRO duality.
\end{distillation}

\section*{Lean comparison targets}
If the project formalizes a generic Gibbs/Donsker--Varadhan lemma, this paper is a friendly secondary source.  For primary citation, prefer Donsker--Varadhan or Dupuis--Ellis if those PDFs are later added.

\section*{Distillation gaps}
\TODO{Do not use this as the primary reference for entropy duality unless a more direct source is unavailable.  It is pedagogical and PAC-Bayes focused.}

\input{references_manual}
\end{document}
