\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{enumitem}
\usepackage[hidelinks]{hyperref}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\newtheorem{remark}{Remark}
\newcommand{\R}{\mathbb{R}}
\newcommand{\e}{\mathrm{e}}
\newcommand{\diam}{\operatorname{diam}}
\newcommand{\spanop}{\operatorname{span}}
\newcommand{\supp}{\operatorname{supp}}

\title{Modernized proof notes: Birkhoff 1957, Extensions of Jentzsch's theorem}
\author{}
\date{2026-07-09}

\begin{document}
\maketitle

\section*{Source map and scope}
\begin{sloppypar}
\begin{description}[style=nextline,leftmargin=3.2cm,labelwidth=3cm]
\item[Primary source] \path{Birkhoff1957_extensions_jentzsch.pdf}
\item[Coverage] The projective metric construction, contraction theorem, fixed-ray theorem, and the Jentzsch-type spectral-gap argument in Sections 2--8.
\item[Source anchors]
\begin{itemize}[leftmargin=2em]
  \item Section 2, equations (2)--(7)
  \item Section 3, equation (8), Lemma 1, and Examples 1--3
  \item Section 4, equations (9)--(9a), Theorem 1, and its two corollaries
  \item Section 5, equations (10)--(12) and Theorem 2
  \item Section 6, equations (13)--(15), Lemma 2, its two corollaries, and Theorem 3
  \item Section 7, equations (16a)--(17), Lemma 3, its corollary, and Theorem 4
  \item Section 8, equations (18)--(22), Theorem 5, and Lemma 4
\end{itemize}
\end{description}
\end{sloppypar}

The exposition below preserves the source proof order and proof technique.  Notation is normalized
and intermediate calculations are made explicit.  When the source contains a gap, ambiguity, or
apparently incorrect step, the reconstruction records the source step before adding a separate
commentary remark.

\section*{Reference being modernized}
Garrett Birkhoff. ``Extensions of Jentzsch's theorem.'' \emph{Transactions of the American Mathematical Society} 85(1):219--227, 1957.
The source is Birkhoff's 1957 Transactions paper, pp. 219--227.

\section*{Purpose of this note}
This reconstruction follows the paper almost section by section, retaining the line-metric calculation, the contraction argument, and the Jentzsch-type spectral conclusion while modernizing notation.

The principal statements reconstructed below are:
\begin{enumerate}[leftmargin=2em]
\item the exact one-dimensional projective-line computation leading to
\[
  N(P)=\tanh(\lambda/4),
\]
where \(\lambda\) is the projective diameter of the image interval;
\item the reduction from a cone to the projective line through two rays;
\item the finite-diameter \(\Rightarrow\) strict-contraction \(\Rightarrow\) unique characteristic ray argument;
\item the positive-matrix specialization: a strictly positive matrix maps the positive simplex into a compact subset of its interior, hence its image has finite projective diameter.
\end{enumerate}

\section{Translation of Birkhoff's notation}
Birkhoff writes vectors on the right, so a linear operator is denoted \(fP\).  We keep that convention when following his proof.  For comparison with modern cone-metric notation, a modern left-acting operator \(Lx\) corresponds to Birkhoff's \(xP\).

Birkhoff writes \(\theta(f,g;C)\) for Hilbert's projective distance.  In current notes we usually write \(d_C(f,g)\).  The following table fixes the dictionary.
\[
\begin{array}{c|c}
\text{Birkhoff 1957} & \text{Modern notation}\\ \hline
fP & Lf\\
\theta(f,g;C) & d_C(f,g)\\
N(P;C) & \kappa(L;C,C)\\
CP & L(C)\\
\text{diameter }A\text{ of }CP & \Delta(L)=\diam_C(L(C))
\end{array}
\]

Birkhoff works with a projectivized cone: a ray is represented by the unique intersection with a hyperplane.  In finite positive cones this is the simplex cross-section \(\sum_i x_i=1\).  The metric is invariant under positive scalar multiplication, so all estimates are statements about rays.

\section{Section 2: the projective-line calculation}

\subsection{Positive quadrant as a projective interval}
Consider the open first quadrant in \(\R^2\).  A positive ray is represented by homogeneous coordinates
\[
  f=(f_1,f_2),\qquad f_1>0,
\]
and by the affine/projective coordinate
\[
  x=\frac{f_2}{f_1}\in(0,\infty).
\]
Birkhoff maps this interval to the real line by
\[
  u=\log x=\log(f_2/f_1).
\]
The projective distance between rays \(f=(f_1,f_2)\) and \(g=(g_1,g_2)\) is therefore
\begin{equation}
  \theta(f,g)=\left|\log\frac{f_2/f_1}{g_2/g_1}\right|
  =\left|\log\frac{f_2g_1}{f_1g_2}\right|.
  \label{eq:birkhoff-line-metric}
\end{equation}
This is the logarithm of the cross-ratio with endpoints \(0\) and \(\infty\):
\[
  \frac{f_2g_1}{f_1g_2}=R(f_2/f_1,g_2/g_1;0,\infty).
\]
Thus any projective transformation preserving the interval \((0,\infty)\) preserves the form of the cross-ratio metric before restriction to a smaller interval.

\subsection{A positive fractional-linear map}
Birkhoff now considers a projective transformation
\begin{equation}
  y=xP=\frac{ax+b}{cx+d}.
  \label{eq:fractional-linear}
\end{equation}
Assume it maps \((0,\infty)\) into a proper subinterval of \((0,\infty)\).  Composing with the isometry \(y\mapsto 1/y\) if necessary, one may assume the map preserves order.  Then its coefficients may be taken positive and
\[
  a,b,c,d>0,\qquad ad>bc.
\]
Here is Birkhoff's sign check in modern form.
\begin{itemize}[leftmargin=2em]
\item Since \(0P=b/d>0\), \(b\) and \(d\) have the same sign.
\item Since \(xP=0\) has no solution with \(x>0\), \(a\) and \(b\) have the same sign.
\item Since \(xP=\infty\) has no solution with \(x>0\), \(c\) and \(d\) have the same sign.
\item Multiplying numerator and denominator by \(-1\), all four coefficients can be chosen positive.
\item Order preservation is \(dy/dx>0\), i.e.
\[
  \frac{dy}{dx}=\frac{ad-bc}{(cx+d)^2}>0,
\]
so \(ad>bc\).
\end{itemize}

\subsection{Differential contraction factor}
On the logarithmic coordinate, the infinitesimal distance is \(d\theta(x)=|d\log x|=|dx|/x\).  For \(y=y(x)\), the differential ratio is
\begin{equation}
  \frac{d\theta(y)}{d\theta(x)}
  =\frac{|dy|/y}{|dx|/x}
  =\frac{x}{y}\frac{dy}{dx}.
\end{equation}
Substitute \eqref{eq:fractional-linear} and its derivative:
\[
  \frac{x}{y}\frac{dy}{dx}
  =x\frac{cx+d}{ax+b}\frac{ad-bc}{(cx+d)^2}
  =\frac{(ad-bc)x}{(ax+b)(cx+d)}.
\]
Define
\begin{equation}
  r(x)=\frac{(ad-bc)x}{(ax+b)(cx+d)}.
  \label{eq:differential-ratio}
\end{equation}
The operator norm on the projective line is the supremum of \(r(x)\) over \(x>0\).  Since the numerator is positive, maximize the reciprocal denominator relative to \(x\), or differentiate.  Write
\[
  (ax+b)(cx+d)=acx^2+(ad+bc)x+bd.
\]
Then
\[
  r'(x)=0
  \quad\Longleftrightarrow\quad
  (ax+b)(cx+d)-x\bigl(a(cx+d)+c(ax+b)\bigr)=0.
\]
Expanding and cancelling gives
\[
  acx^2+(ad+bc)x+bd-x(2acx+ad+bc)=0,
\]
hence
\[
  bd-acx^2=0,
  \qquad x_*=\sqrt{\frac{bd}{ac}}.
\]
At this point,
\[
  ax_* = a\sqrt{\frac{bd}{ac}}=\sqrt{\frac{abd}{c}},
  \qquad cx_* = c\sqrt{\frac{bd}{ac}}=\sqrt{\frac{bcd}{a}}.
\]
A cleaner simplification uses the projective invariant
\[
  v=\frac{ad}{bc}>1.
\]
At \(x=x_*\), one obtains
\begin{equation}
  N(P)=\sup_{x>0}r(x)=\frac{v-1}{v+2\sqrt v+1}
       =\frac{\sqrt v-1}{\sqrt v+1}.
  \label{eq:N-line-v}
\end{equation}
This is Birkhoff's formula (5), rewritten in the equivalent simplified form.

\subsection{Composition submultiplicativity}
Because \(N(P)\) is defined as the supremum of all distance ratios,
\[
  N(P)=\sup_{f,g}\frac{\theta(fP,gP)}{\theta(f,g)},
\]
we immediately have
\begin{equation}
  N(PP')\le N(P)N(P').
  \label{eq:N-submultiplicative}
\end{equation}
Indeed,
\[
  \frac{\theta(fPP',gPP')}{\theta(f,g)}
  =\frac{\theta((fP)P',(gP)P')}{\theta(fP,gP)}
   \frac{\theta(fP,gP)}{\theta(f,g)}
  \le N(P')N(P),
\]
whenever denominators are nonzero; degenerate cases follow by continuity or direct convention.

\subsection{Image length and the hyperbolic tangent}
The map sends the endpoints \(0,\infty\) to
\[
  P(0)=\frac{b}{d},\qquad P(\infty)=\frac{a}{c}.
\]
Thus the length of the image interval in the logarithmic metric is
\begin{equation}
  \lambda=\log\frac{a/c}{b/d}=\log\frac{ad}{bc}=\log v.
\end{equation}
Therefore \(v=\e^\lambda\), \(\sqrt v=\e^{\lambda/2}\), and \eqref{eq:N-line-v} becomes
\[
  N(P)=\frac{\e^{\lambda/2}-1}{\e^{\lambda/2}+1}
       =\frac{\e^{\lambda/4}-\e^{-\lambda/4}}{\e^{\lambda/4}+\e^{-\lambda/4}}
       =\tanh(\lambda/4).
\]
This is the exact source of the \(1/4\) in the Birkhoff--Hopf contraction coefficient.

\section{Section 3: convex cones and the associated projective metric}

\subsection{Projectivized cone}
Let \(C\) be a closed convex cone in a real vector space \(L\).  Birkhoff assumes that one can choose a hyperplane \(H\) meeting each ray of \(C\) in exactly one point.  Then the cone can be identified with the convex cross-section
\[
  C\cap H.
\]
In finite positive cones this is the standard simplex.  In Banach-space examples this is often a normalized positive part such as \(\|f\|=1\) or \(\int f=1\), whenever it is available.

Given two distinct points \(f,g\in C\cap H\), the affine line through them intersects the closed convex set \(C\cap H\) in a closed segment with endpoints \(a,b\).  There is a unique affine coordinate on this line carrying \(a\) to \(0\), \(b\) to \(\infty\), and the interior of the segment to \((0,\infty)\).  Birkhoff defines
\begin{equation}
  \theta(f,g;C)=\theta(fA,gA)
  \label{eq:theta-cone}
\end{equation}
where \(A\) is that affine/projective coordinate and the right-hand side is the line metric \eqref{eq:birkhoff-line-metric}.  If either point lies on the boundary, Birkhoff sets the distance to \(\infty\) in this formulation.

Modern order-unit notation gives the same distance for comparable interior points:
\[
  d_C(x,y)=\log\frac{M(x/y)}{m(x/y)},
\]
where
\[
  M(x/y)=\inf\{\beta:x\le_C \beta y\},
  \qquad
  m(x/y)=\sup\{\alpha:\alpha y\le_C x\}.
\]

\begin{lemma}[Birkhoff's metric-space lemma]
Fix a ray \(a\in C\).  Let \(A\) be the set of rays \(f\in C\) such that \(\theta(a,f;C)<\infty\).  Then \(A\) is a metric space under \(\theta(\cdot,\cdot;C)\) after identifying scalar multiples.
\end{lemma}

\begin{proof}[Expansion of a proof omitted by the source]
Birkhoff calls this fact well known and gives no proof.  The following order-theoretic expansion, in the standard modern formulation used by
Lemmens--Nussbaum~\cite{LemmensNussbaum2013}, supplies the cited Hilbert-metric argument without
changing any later source step.
The projective line through any two comparable rays cuts the cone in a segment, so the distance between the two rays is the logarithm of the corresponding cross-ratio.  Positivity is immediate from the ordering of the four points on the segment.  Symmetry follows because the cross-ratio is inverted when the two interior points are interchanged, and the logarithm is taken in absolute value.  The triangle inequality is Hilbert's standard triangle inequality on a convex domain; in order notation it follows from
\[
  M(x/z)\le M(x/y)M(y/z),
  \qquad
  m(x/z)\ge m(x/y)m(y/z).
\]
Indeed, if \(x\le M(x/y)y\) and \(y\le M(y/z)z\), then \(x\le M(x/y)M(y/z)z\).  Similarly, if \(m(x/y)y\le x\) and \(m(y/z)z\le y\), then \(m(x/y)m(y/z)z\le x\).  Taking logarithms gives
\[
  d_C(x,z)\le d_C(x,y)+d_C(y,z).
\]
Zero distance identifies rays in the usual closed proper cone setting; Birkhoff is already working projectively, so this is exactly the metric assertion on rays.
\end{proof}

\subsection{Examples used by Birkhoff}
Birkhoff lists three examples.

\paragraph{Example 1: finite positive orthant.}
Let
\[
  R_+^n=\{f\in\R^n:f_i\ge0\},
  \qquad H=\{f:\sum_i f_i=1\}.
\]
The cross-section is the closed simplex.  The components for the projective metric are the relative interiors of faces.  On the strictly positive part,
\[
  d(f,g)=\log\frac{\max_i f_i/g_i}{\min_i f_i/g_i}.
\]
This is the finite-coordinate cross-ratio formula.

\paragraph{Example 2: continuous positive functions.}
Let \(L=C(S)\) with the Banach lattice norm \(\|f\|=\sup_x |f(x)|\), and let \(C=L_+\) be the cone of nonnegative functions.  The strictly positive functions form a component.  In that component,
\[
  d(f,g)=\log\frac{\sup_x f(x)/g(x)}{\inf_x f(x)/g(x)}.
\]

\paragraph{Example 3: bounded measurable functions with bounded oscillation ratio.}
Let \(B\) be the Banach space of bounded measurable functions on the unit interval.  For fixed \(M>0\), Birkhoff considers the cone
\[
  C_M=\{f:0<\sup f\le M\inf f\}\cup\{0\}.
\]
This is adapted to integro-functional operators that preserve a uniform upper/lower bound ratio.

\section{Section 4: the projective contraction theorem}
Let \(P\) be a bounded linear map of a Banach space into itself with \(P(C)\subseteq C\).  Birkhoff defines the projective norm
\begin{equation}
  N(P;C)=\sup\left\{\frac{\theta(fP,gP;C)}{\theta(f,g;C)}:
        f,g\in C,\;\theta(f,g;C)<\infty\right\}.
\end{equation}

\begin{lemma}[Birkhoff's finite-diameter contraction lemma]
If \(CP=P(C)\) has finite projective diameter \(A\) in \((C,\theta)\), then
\begin{equation}
  N(P;C)=\tanh(A/4)<1.
  \label{eq:birkhoff-diameter-contraction}
\end{equation}
If \(CP\) has infinite projective diameter, then \(N(P;C)=1\).
\end{lemma}

\begin{proof}
Take \(f,g\in C\) with finite \(\theta(f,g;C)\).  The projective line segment through their rays is a cone interval \(s(a,b)\subseteq C\).  Under \(P\), this interval maps into the projective segment joining \(aP\) and \(bP\), and that segment is contained in \(CP\).  Hence
\[
  \theta(aP,bP;C)\le A.
\]
On this two-dimensional projective line, Section 2 applies.  A fractional-linear map whose image interval has length at most \(A\) contracts distances by at most \(\tanh(A/4)\).  Therefore
\[
  \theta(fP,gP;C)
  \le \tanh(A/4)\theta(f,g;C).
\]
Taking the supremum over \((f,g)\) gives \(N(P;C)\le \tanh(A/4)\).

For the reverse inequality, choose image pairs on line segments in \(CP\) whose projective distance tends to \(A\).  Pull those points back to preimage points \(f_n,g_n\in C\).  The one-dimensional calculation on the corresponding segment says that the best possible contraction ratio on that segment is \(\tanh(\lambda_n/4)\), where \(\lambda_n\to A\).  Hence
\[
  N(P;C)\ge \lim_n \tanh(\lambda_n/4)=\tanh(A/4).
\]
If \(A=\infty\), the same line-wise calculation gives ratios arbitrarily close to \(1\), while nonexpansiveness gives \(N(P;C)\le1\).  Thus \(N(P;C)=1\).
\end{proof}

\begin{theorem}[Birkhoff's projective contraction theorem]
Assume \(N(P^r;C)<1\) for some integer \(r\ge1\), and assume the relevant projective component of \(C\) is complete under \(\theta\).  Then for every \(f\in C\) in that component, the sequence of rays \(fP^n\) converges geometrically to a unique fixed ray \(c\), i.e. a characteristic ray satisfying
\[
  cP=\gamma c
\]
for some scalar \(\gamma>0\).
\end{theorem}

\begin{proof}
Let \(\rho=N(P^r;C)<1\).  Since \(P^r(C)\) has finite diameter, the distance
\[
  \theta(fP^r,fP^{r+1};C)
\]
is finite.  More generally, write \(q=\lfloor n/r\rfloor\).  Applying the contraction estimate for \(P^r\) repeatedly gives
\begin{equation}
  \theta(fP^n,fP^{n+1};C)
  \le \rho^{q-1}\theta(fP^r,fP^{r+1};C).
  \label{eq:successive-distance-geometric}
\end{equation}
The right-hand side is summable in \(n\).  Hence, for \(m>n\), the triangle inequality gives
\[
  \theta(fP^n,fP^m;C)
  \le \sum_{k=n}^{m-1}\theta(fP^k,fP^{k+1};C)\longrightarrow0.
\]
Thus \(\{fP^n\}\) is Cauchy in projective metric.  Completeness gives a projective limit ray \(c\in C\).

Since \(P\) is bounded, the normalized representatives can be chosen so that applying \(P\) commutes with the limit.  Therefore the image of the limit ray is the limit of the shifted sequence:
\[
  cP=\lim_n fP^{n+1}.
\]
The shifted sequence has the same projective limit as \(fP^n\), because the successive distance tends to zero by \eqref{eq:successive-distance-geometric}.  Hence \(cP\) lies on the same ray as \(c\): there exists \(\gamma>0\) with \(cP=\gamma c\).

For uniqueness, suppose \(c\) and \(c^*\) are fixed rays for \(P\).  Then they are fixed rays for \(P^r\), so
\[
  \theta(c,c^*)=\theta(cP^r,c^*P^r)
  \le N(P^r;C)\theta(c,c^*)
  =\rho\theta(c,c^*).
\]
Since \(\rho<1\), the only finite possibility is \(\theta(c,c^*)=0\), so the rays agree.
\end{proof}

\begin{corollary}
If \(C P^r\) has finite projective diameter for some \(r\), then the conclusion of the projective contraction theorem holds.
\end{corollary}

\section{Section 5: applications}

\subsection{Positive matrices}
Let \(P=(p_{ij})\) be a matrix with strictly positive entries.  Acting on the positive cone \(R_+^n\), it maps the simplex
\[
  \Sigma=\{x\ge0:\sum_i x_i=1\}
\]
into a compact subset of the simplex interior.  Equivalently, for every \(i\),
\[
  (xP)_i=\sum_j x_jp_{ji}>0
\]
for nonzero \(x\ge0\), and compactness of \(\Sigma\) plus positivity of all entries gives uniform positive lower and upper coordinate bounds on the normalized image.

Thus \(P(\Sigma)\) has finite projective diameter.  The contraction theorem gives a unique positive eigenvector ray \(c\in R_+^n\), with
\[
  cP=\gamma c,
\]
and iterates of every positive ray converge projectively to \(c\).  Birkhoff notes that it suffices for \(P\) to be nonnegative and for some power \(P^r\) to be strictly positive.

In finite dimensions, the compactness sentence can be replaced by the explicit matrix diameter bound.  If \(A=(a_{ij})\) is strictly positive, then for \(x,y>0\),
\[
  d(Ax,Ay)\le \log\max_{p,q,i,j}\frac{a_{pi}a_{qj}}{a_{pj}a_{qi}}.
\]
This bound is finite because the index sets are finite and all entries are positive.

\subsection{Jentzsch-type integral operators}
Birkhoff next considers
\[
  [fP](x)=\int_0^1 p(x,y)f(y)\,dy,
\]
where the kernel satisfies uniform two-sided bounds
\begin{equation}
  0<\gamma=\inf_{x,y}p(x,y)\le \sup_{x,y}p(x,y)=K\gamma=S.
  \label{eq:birkhoff-kernel-bounds}
\end{equation}
Let \(e(x)=1\).  If \(f\ge0\) and
\[
  \varphi=\int_0^1 f(y)\,dy>0,
\]
then \eqref{eq:birkhoff-kernel-bounds} gives
\[
  \gamma\varphi\,e\le fP\le S\varphi\,e.
\]
Therefore the projective distance from \(fP\) to \(e\) is bounded by
\[
  \theta(e,fP;L_+)\le \log(S/\gamma)=\log K.
\]
Consequently \(P(L_+)\) has finite projective diameter, and the contraction theorem produces a unique positive characteristic function \(c\) satisfying
\[
  cP=\gamma_0 c
\]
for some scalar \(\gamma_0>0\).  This recovers the essence of Jentzsch's theorem without Fredholm theory.

\subsection{Integro-functional equation}
Birkhoff's Theorem 2 treats
\begin{equation}
  [fP](x)=\int_0^1 p(x,y)f(y)\,dy+
          \sum_{i=1}^n a_i f(xT_i),
  \label{eq:integro-functional}
\end{equation}
where \(p\) satisfies \eqref{eq:birkhoff-kernel-bounds}, the \(T_i\) are one-to-one Borel maps of the unit interval onto itself, and \(a_i>0\) with \(A=\sum_i a_i\).  Choose \(M>K\), and use the cone
\[
  C_M=\{f:0<\sup f\le M\inf f\}\cup\{0\}.
\]
For \(f\in C_M\), set \(g=fP\).  Then
\[
  \sup_x g(x)
  \le S\int f(y)\,dy+MA\inf_x f(x),
\]
while
\[
  \inf_x g(x)
  \ge \gamma\int f(y)\,dy+A\inf_x f(x).
\]
Using \(\int f\ge \inf f\), Birkhoff obtains
\begin{equation}
  \frac{\sup g}{\inf g}
  \le \frac{K\int f+MA\inf f}{\int f+A\inf f}
  \le M,
  \label{eq:integro-functional-cone-invariance}
\end{equation}
provided \(K<M\).  Thus \(P(C_M)\subseteq C_M\).  Moreover the same estimate gives a bounded projective diameter of the image, so the contraction theorem yields a unique positive characteristic function \(c\) satisfying
\[
  \int_0^1 p(x,y)c(y)\,dy+
  \sum_{i=1}^n a_i c(xT_i)=\gamma c(x).
\]

\section{Section 6: Banach lattices and uniform positivity}

\subsection{Strong comparability in a vector lattice}
Let \(L\) be a vector lattice and \(L_+\) its positive cone.  Birkhoff's Lemma 2 says that if \(f,g\) are in the same projective component of \(L_+\), then they are strongly comparable: there are scalars \(\lambda,\mu>0\) and a finite \(R\) such that
\begin{equation}
  \lambda f\le g\le R\lambda f,
  \qquad
  \mu g\le f\le R\mu g.
  \label{eq:strong-comparable}
\end{equation}
The smallest such \(R\) is
\[
  R=\exp(\theta(f,g;L_+)).
\]

The proof is two-dimensional.  The plane \(\spanop\{f,g\}\) intersects \(L_+\) in a cone affine-equivalent to a quadrant.  In quadrant coordinates, if
\[
  f=(f_1,f_2),\qquad g=(g_1,g_2),
\]
then comparability gives positive coordinate ratios, and
\[
  R=\max\left\{\frac{g_1/f_1}{g_2/f_2},\frac{g_2/f_2}{g_1/f_1}\right\}
  =\exp\left(\left|\log\frac{f_2g_1}{f_1g_2}\right|\right).
\]
This is exactly the exponential of the projective distance.

\subsection{Norm consequence}
On the unit sphere of a Banach lattice,
\begin{equation}
  \|f-g\|\le \e^\theta-1,
  \qquad \theta=\theta(f,g;L_+).
  \label{eq:norm-projective-control}
\end{equation}
Indeed, from \eqref{eq:strong-comparable} with \(R=\e^\theta\) and \(\|f\|=\|g\|=1\), monotonicity of the lattice norm gives \(\lambda\le1\).  Then the lattice positive and negative parts satisfy
\[
  |f-g|\le (R-1)\lambda |f|,
\]
so
\[
  \|f-g\|\le (R-1)\lambda\|f\|
  \le \e^\theta-1.
\]
Consequently each projective component of the unit sphere is complete if the Banach lattice is norm complete.

\subsection{Uniform positivity}
Birkhoff calls a bounded linear map \(P:L\to L\) uniformly positive if there exist a fixed \(e>0\) in \(L\) and a finite constant \(K\) such that for each \(f>0\) there is a scalar \(\lambda=\lambda(f)>0\) with
\begin{equation}
  \lambda e\le fP\le K\lambda e.
  \label{eq:uniform-positive}
\end{equation}
This says exactly that every image ray lies within projective distance at most \(\log K\) from the ray of \(e\).  Therefore \(P(L_+)\) has projective diameter at most \(\log K\).

\begin{theorem}[Birkhoff's uniformly positive Banach-lattice theorem]
A uniformly positive bounded linear map \(P\) on a Banach lattice \(L\) admits a unique positive unit eigenvector \(c\) and positive eigenvalue \(\gamma\):
\[
  cP=\gamma c.
\]
Moreover for every \(f>0\), normalized iterates converge geometrically:
\[
  \left\|\frac{fP^n}{\|fP^n\|}-c\right\| \le M\rho^n
\]
for some \(M<\infty\) and \(0<\rho<1\).
\end{theorem}

\begin{proof}
Use $C=L_+$.  By the completeness corollary to Lemma 2, it remains only to prove that $CP$ has
finite projective diameter.  For $f,g>0$, choose the scalars in the uniform-positivity condition so
that
\[
 e\le fP\le Ke,
 \qquad
 e\le gP\le Ke.
\]
The four positive elements
\[
 KfP-gP,\qquad (K-1)fP,\qquad (K-1)gP,\qquad KgP-fP
\]
lie in the cone.  On the two-dimensional projective line through $fP$ and $gP$, these four points
are perspective to the scalar quadruple
\[
 -1,\quad 0,\quad K-1,\quad K.
\]
Projective invariance of the cross-ratio therefore gives
\[
 \theta(fP,gP;L_+)\le\log K.
\]
Thus $CP$ has diameter at most $\log K$.  Theorem 1 yields the unique characteristic ray and
geometric convergence in projective distance.  Equation \eqref{eq:norm-projective-control} converts
that convergence, after normalization to the unit sphere, into the asserted Banach-lattice norm
convergence.
\end{proof}

\section{Section 7: complementary invariant subspace}
Let $P$ be uniformly positive, let $c>0$ be its normalized characteristic vector, and write
\[
 cP=\gamma c,
 \qquad \gamma>0.
\]
For $f>0$, Birkhoff defines $\lambda_n$ and $\mu_n$ as the largest and smallest scalars for which
\begin{equation}
 \lambda_n\gamma^n c\le fP^n\le\mu_n\gamma^n c.
 \tag{16a}
\end{equation}
Applying $P$ and using $cP=\gamma c$ gives
\begin{equation}
 \lambda_n\le\lambda_{n+1}\le\mu_{n+1}\le\mu_n.
 \tag{16b}
\end{equation}
Thus $(\lambda_n)$ increases and $(\mu_n)$ decreases.

Set
\[
 r_n=fP^n-\lambda_n\gamma^n c,
 \qquad
 s_n=\mu_n\gamma^n c-fP^n.
\]
Then $r_n,s_n\ge0$ and
\[
 r_n+s_n=(\mu_n-\lambda_n)\gamma^n c.
\]
Uniform positivity, together with Lemma 2, gives $\alpha_n,\beta_n>0$ and a projective diameter
bound $A$ such that
\begin{equation}
 \alpha_nc\le r_nP\le e^A\alpha_nc,
 \qquad
 \beta_nc\le s_nP\le e^A\beta_nc.
 \tag{16c}
\end{equation}
Adding the inequalities and using $(r_n+s_n)P=(\mu_n-\lambda_n)\gamma^{n+1}c$ yields
\begin{equation}
 (\alpha_n+\beta_n)c
 \le(\mu_n-\lambda_n)\gamma^{n+1}c
 \le e^A(\alpha_n+\beta_n)c.
 \tag{16d}
\end{equation}
On the other hand,
\begin{align*}
 fP^{n+1}
 &=\lambda_n\gamma^{n+1}c+r_nP
 \ge(\lambda_n\gamma^{n+1}+\alpha_n)c,\\
 fP^{n+1}
 &=\mu_n\gamma^{n+1}c-s_nP
 \le(\mu_n\gamma^{n+1}-\beta_n)c.
\end{align*}
By extremality of $\lambda_{n+1}$ and $\mu_{n+1}$,
\[
 \lambda_{n+1}\ge\lambda_n+\gamma^{-n-1}\alpha_n,
 \qquad
 \mu_{n+1}\le\mu_n-\gamma^{-n-1}\beta_n.
\]
Subtracting and applying the right-hand inequality in (16d) gives
\[
 \mu_{n+1}-\lambda_{n+1}
 \le(1-e^{-A})(\mu_n-\lambda_n).
\]
Hence both monotone sequences converge to a common limit $M(f)$, and
\[
 \mu_n-\lambda_n
 \le(1-e^{-A})^{n-1}(\mu_1-\lambda_1).
\]
Returning to (16a) gives Birkhoff's Lemma 3: for each $f>0$ there are constants $M(f)>0$ and
$K(f)<\infty$ such that
\begin{equation}
 |fP^n-M(f)\gamma^n c|\le K(f)\rho^n c,
 \qquad
 \rho=(1-e^{-A})\gamma<\gamma.
 \tag{17}
\end{equation}

Every $f\in L$ has the lattice decomposition $f=f^+-f^-$.  Applying (17) to $f^+$ and $f^-$
extends $M$ to a positive linear functional on $L$.  Let
\[
 S=\ker M.
\]
Equation (17) implies $SP\subseteq S$.  Moreover, on $S$ the spectral radius of $P$ is at most
$(1-e^{-A})\gamma<\gamma$.  This is Theorem 4: the space decomposes into the leading positive axis
$\R c$ and a complementary invariant subspace with strictly smaller spectral growth.

\begin{remark}
The source calls the bound in Theorem 4 the ``spectral norm.''  In modern terminology the proof
controls the asymptotic growth/spectral radius on the invariant complement; it does not construct an
orthogonal decomposition or claim that $P$ is normal.
\end{remark}

\section{Section 8: multiplicative processes}
Let $(P_t)_{t>0}$ be a semigroup of nonnegative bounded linear maps on a Banach lattice:
\begin{equation}
 f>0\Longrightarrow fP_t>0,
 \qquad
 (fP_t)P_s=fP_{t+s}. \tag{18}
\end{equation}
The motivating kernel form is
\begin{equation}
 f(x;t+\tau)=\int p(x,y;\tau)f(y;t)\,dR(y),
 \qquad p(x,y;\tau)>0. \tag{19}
\end{equation}
Assume $P_T$ is uniformly positive for some $T>0$.  Apply Theorem 4 to the discrete operator
$P_T$.  There are $c>0$, $\gamma=e^{\beta T}>0$, a positive linear functional $m$, and
$0\le\rho<\gamma$ such that
\[
 \|fP_{nT}-m(f)\gamma^n c\|\le K(f)\rho^n.
\]

For $t>nT$, semigroup monotonicity of image diameter gives
\begin{equation}
 \Delta(CP_t)
 =\Delta((CP_{t-nT})P_{nT})
 \le\Delta(CP_{nT})\longrightarrow0. \tag{21}
\end{equation}
Therefore the characteristic ray selected by every $P_t$ is the same ray $c$.  The corresponding
multiplicative eigenvalues form a positive scalar semigroup, hence are $e^{\beta t}$ for a unique
real growth coefficient $\beta$.

Write $f=m(f)c+r$ with $r\in\ker m$.  For $t=nT+s$ with $s$ in a bounded interval,
\begin{align*}
 \|fP_t-e^{\beta t}m(f)c\|
 &=\|(fP_{nT}-m(f)\gamma^nc)P_s\|\\
 &\le K(f)\rho^n\|P_s\|.
\end{align*}
The source obtains uniform boundedness of $\|P_s\|$ for $s$ in the required compact interval from
the order bounds underlying (16d).  Choose $\sigma<\beta$ with $e^{\sigma T}=\rho$.  Absorbing the
bounded remainder interval into the constant gives Theorem 5:
\begin{equation}
 \|fP_t-e^{\beta t}m(f)c\|
 \le K^*(f)e^{\sigma t},
 \qquad 0\le\sigma<\beta. \tag{20,22}
\end{equation}

The final Lemma 4 extends the completeness input from Banach lattices to complete vector lattices.
From a projective Cauchy sequence one extracts a subsequence $(f_n)$ with
$\theta(f_n,f_{n+1})<2^{-n}$.  Strong comparability gives
\[
 |f_{n+1}-f_n|
 \le(e^{2^{-n}}-1)|f_n|<2^{-n}|f_n|.
\]
Induction bounds all $f_n$ by a common order interval.  Order completeness then produces the
supremum/limit, and the same estimates show relative-uniform and intrinsic convergence.  Hence
each projective component is complete.

\section{Source-faithfulness notes}
\begin{itemize}[leftmargin=2em]
\item Birkhoff does not prove the general Hilbert-metric triangle inequality in Section 3; he calls
it well known and cites projective-geometry sources.  The explicit order proof above is additional
exposition, not a source step.
\item Equality in the finite-diameter contraction formula is justified in the source by taking
preimages of nearby image pairs on segments whose lengths approach the diameter.  The argument is
brief and presupposes that the supremal diameter can be approximated in this way.
\item In Theorem 1 the source writes $cP=c$ after passing to projective representatives.  The
mathematical conclusion is a fixed ray, equivalently $cP=\gamma c$ before normalization.
\item Theorem 2's printed estimate contains typography that is difficult to parse in scanned copies.
The invariant-cone conclusion uses exactly the displayed upper and lower estimates and the choice
$M>K$; no sharper ratio is needed.
\item The source's use of ``spectral norm'' in Theorem 4 predates current operator-norm terminology.
The proof is an asymptotic spectral-gap estimate on $\ker M$.
\end{itemize}

\section{Finite-dimensional specialization and relation to Sinkhorn iteration}
The finite-dimensional core of the argument is:
\begin{enumerate}[leftmargin=2em]
\item Define the finite Hilbert metric on strictly positive vectors by
\[
  d_H(x,y)=\log\frac{\max_i x_i/y_i}{\min_i x_i/y_i}.
\]
\item Prove the two-dimensional line calculation or import it as the theorem seam:
\[
  \kappa(L)=\tanh(\Delta(L)/4).
\]
\item Prove strict positivity of a finite matrix gives finite projective image diameter, preferably by the explicit cross-ratio bound.
\item Use projective contraction/nonexpansiveness to obtain uniqueness of cluster rays or asymptotic regularity.
\end{enumerate}

The original Birkhoff paper does not prove Sinkhorn's alternating nonlinear scaling iteration.  It supplies the projective contraction tool and the positive-matrix fixed-ray mechanism.  Eveson--Nussbaum give a later elementary proof of the Birkhoff--Hopf theorem~\cite{EvesonNussbaum1995}, while Franklin--Lorenz apply projective contraction to matrix scaling~\cite{FranklinLorenz1989}.  Those later arguments are supplementary references, not steps attributed to Birkhoff.

\begin{thebibliography}{9}
\bibitem{Birkhoff1957} G. Birkhoff, \emph{Extensions of Jentzsch's theorem}, Trans. Amer. Math. Soc. 85(1):219--227, 1957.
\bibitem{LemmensNussbaum2013} B. Lemmens and R. Nussbaum, \emph{Birkhoff's version of Hilbert's metric and its applications in analysis}, arXiv:1304.7921, 2013.
\bibitem{EvesonNussbaum1995} S. P. Eveson and R. D. Nussbaum, \emph{An elementary proof of the Birkhoff--Hopf theorem}, Math. Proc. Camb. Phil. Soc. 117:31--55, 1995.
\bibitem{FranklinLorenz1989} J. Franklin and J. Lorenz, \emph{On the scaling of multidimensional matrices}, Linear Algebra Appl. 114/115:717--735, 1989.
\end{thebibliography}

\end{document}
