\input{preamble}
\title{Strong Duality for Distributional Model Risk under Optimal Transport}
\author{}
\date{Blanchet and Murthy (2019), Theorem 1 and its proof}
\begin{document}
\maketitle

\section*{Source map and scope}
\begin{description}[style=nextline,leftmargin=3.2cm,labelwidth=3cm]
\item[Primary source] \texttt{\detokenize{blanchet-murthy-2019_quantifying-distributional-model-risk-OT__arXiv-1604.01446.pdf}}
\item[Coverage] A line-by-line reconstruction of Theorem 1(a)--(b), including the compact minimax reduction, exhaustion of the general Polish space, measurable selection, and complementary slackness.
\item[Source anchors]
\begin{itemize}[leftmargin=2em]
  \item Theorem 1
  \item equations (7)--(9)
  \item Propositions 5--7
  \item Lemma 8
  \item proof of Theorem 1(b)
\end{itemize}
\end{description}

The exposition below preserves the source proof order and proof technique.  Notation is normalized
and intermediate calculations are made explicit.  When the source contains a gap, ambiguity, or
apparently incorrect step, the reconstruction records the source step before adding a separate
commentary remark.

\section{Setting}
Let $S$ be a Polish space, let $\mu\in\Pcal(S)$, and let $\delta>0$.  The transportation cost
$c:S\times S\to[0,\infty)$ and the payoff $f:S\to\R$ satisfy:
\begin{enumerate}[label=(A\arabic*)]
\item $c$ is lower semicontinuous, $c(x,y)=0$ exactly when $x=y$;
\item $f$ is upper semicontinuous and $f\in L^1(\mu)$.
\end{enumerate}
For probability measures $\mu_1,\mu_2$ define
\[
 d_c(\mu_1,\mu_2)
 :=\inf_{\pi\in\Pi(\mu_1,\mu_2)}\int c(x,y)\,\pi(dx,dy).
\]
Equivalently, optimize directly over couplings having first marginal $\mu$:
\[
 \Phi_{\mu,\delta}
 :=\left\{\pi\in\Pcal(S\times S):\pi_1=\mu,
        \ \int c\,d\pi\le\delta\right\},
 \qquad
 I:=\sup_{\pi\in\Phi_{\mu,\delta}}\int f(y)\,d\pi(x,y).
\]
A dual pair is a scalar $\lambda\ge0$ and a universally measurable function $\varphi$ satisfying
\[
 f(y)-\lambda c(x,y)\le\varphi(x),\qquad x,y\in S.
\]
Let
\[
 J:=\inf_{(\lambda,\varphi)}\left\{\lambda\delta+\int\varphi\,d\mu\right\}.
\]
For fixed $\lambda$ the least admissible majorant is the $c$-transform
\[
 \varphi_\lambda(x):=\sup_{y\in S}\{f(y)-\lambda c(x,y)\}.
\]

\section{The strong-duality theorem}
\begin{theorem}[Blanchet--Murthy]
Under the assumptions above:
\begin{enumerate}[label=(\alph*)]
\item $I=J$;
\item the dual infimum is attained by $(\lambda^*,\varphi_{\lambda^*})$ for some $\lambda^*\ge0$;
\item a feasible coupling $\pi^*$ and a dual optimizer $(\lambda^*,\varphi_{\lambda^*})$ are optimal exactly when
\begin{align}
 f(y)-\lambda^*c(x,y)&=\varphi_{\lambda^*}(x)
       &&\pi^*\text{-a.s.},\label{eq:BM-contact}\\
 \lambda^*\left(\int c\,d\pi^*-\delta\right)&=0.\label{eq:BM-slack}
\end{align}
Consequently,
\begin{equation}\label{eq:BM-univariate}
 I=\inf_{\lambda\ge0}
 \left\{\lambda\delta+\int
       \sup_{y\in S}\bigl(f(y)-\lambda c(x,y)\bigr)\,\mu(dx)\right\}.
\end{equation}
\end{enumerate}
\end{theorem}

\section{Weak duality and the least majorant}
For every feasible $\pi$ and every dual pair,
\[
 \int f(y)\,d\pi
 \le \int\varphi(x)\,d\pi+\lambda\int c\,d\pi
 \le \int\varphi\,d\mu+\lambda\delta.
\]
Thus $I\le J$.  Moreover every admissible $\varphi$ dominates $\varphi_\lambda$ pointwise, so once
$\varphi_\lambda$ is known to be universally measurable, the two-variable dual reduces to
\eqref{eq:BM-univariate}.

\section{The compact continuous-cost case}
Assume first that $S$ is compact and $c$ is continuous.  Put $X=C(S\times S)$ and identify
$X^*$ with finite signed Borel measures.  Introduce convex functionals whose common variable is a
continuous test function on $S\times S$:
\begin{align*}
 \Theta(u)&:=
 \begin{cases}
 0,&u(x,y)\ge f(y)\text{ for all }x,y,\\
 +\infty,&\text{otherwise},
 \end{cases}\\
 \Xi(u)&:=
 \begin{cases}
 \displaystyle\int\varphi\,d\mu+\lambda\delta,
 &u(x,y)=\varphi(x)+\lambda c(x,y),\quad \lambda\ge0,\\
 +\infty,&\text{otherwise}.
 \end{cases}
\end{align*}
The Fenchel duality theorem applies because one functional is continuous at a point in the effective
domain of the other.  Computing conjugates gives:
\begin{itemize}[leftmargin=2em]
\item the dual variable must be a nonnegative measure $\pi$;
\item finiteness in the $\varphi$ direction forces $\pi_1=\mu$;
\item finiteness in the $\lambda$ direction forces $\int c\,d\pi\le\delta$;
\item the resulting dual objective is $\int f(y)\,d\pi$.
\end{itemize}
Hence the Fenchel identity is precisely $I=J$.  Compactness of the feasible set and upper
semicontinuity of $\pi\mapsto\int f(y)\,d\pi$ also yield a maximizing coupling.

\section{Compact spaces and lower semicontinuous costs}
On a compact metric space, a nonnegative lower semicontinuous function is the pointwise increasing
limit of continuous functions $c_n$.  Define the corresponding primal and dual values $I_n,J_n$.
The continuous-cost result gives $I_n=J_n$.

Because $c_n\uparrow c$, every $c$-feasible coupling is $c_n$-feasible, so $I\le I_n$.  Conversely,
choose nearly optimal $c_n$-feasible couplings.  Compactness gives a weakly convergent subnet or
subsequence.  Lower semicontinuity and monotone approximation imply that any limit coupling is
$c$-feasible, while upper semicontinuity of $f$ gives the required upper bound on the limiting
payoff.  Therefore $I_n\downarrow I$.  The dual envelopes also decrease to the envelope for $c$,
and monotone convergence gives $J_n\downarrow J$.  Thus $I=J$ for compact $S$ and lower
semicontinuous $c$.

\section{Localization on a general Polish space}
The family $\Phi_{\mu,\delta}$ need not be tight because its second marginals are not fixed.  The
proof therefore localizes the maximization rather than invoking compactness of the whole feasible
set.

Choose compact sets $K_n\uparrow S$ carrying asymptotically all of $\mu$ and on which the relevant
integrals are controlled.  For any finite collection of candidate couplings, enlarge $K_n$ so that the
supports needed by those couplings are contained in a common $\sigma$-compact set.  On such a
localized support the compact-space duality theorem applies after one-point compactification and
truncation.  The resulting statement is the following finite-family reduction: for every finite convex
combination of localized candidates and every error tolerance, there is a feasible coupling whose
payoff is within that tolerance of the corresponding localized robust envelope.

The localized couplings form a convex set $E$.  For a coupling parameter $\pi\in E$ let $S_\pi$ be
its localized second-coordinate support and define
\[
 T(\lambda,\pi)
 :=\lambda\delta+
   \int\sup_{y\in S_\pi}\{f(y)-\lambda c(x,y)\}\,\mu(dx).
\]
For fixed $\pi$, $T$ is convex and lower semicontinuous in $\lambda$; for fixed $\lambda$, its
localized construction is concave in the coupling parameter.  An a priori upper bound on the useful
multipliers restricts $\lambda$ to a compact interval $[0,L]$.  Sion's minimax theorem then gives
\[
 \inf_{0\le\lambda\le L}\sup_{\pi\in E}T(\lambda,\pi)
 =\sup_{\pi\in E}\inf_{0\le\lambda\le L}T(\lambda,\pi).
\]
The compact localized duality result identifies the inner infima with attainable localized primal
values.  Letting the localization expand and the errors vanish gives $J\le I$.  Together with weak
duality this proves $I=J$ on the original Polish space.

\section{Measurability and near-maximizing selectors}
For fixed $\lambda$, the function
\[
 (x,y)\longmapsto f(y)-\lambda c(x,y)
\]
is upper semicontinuous.  Its pointwise supremum over $y$ is upper semianalytic, hence universally
measurable.  More concretely, for each $n$ the set
\[
 A_n:=\left\{(x,y):f(y)-\lambda c(x,y)
                 \ge\varphi_\lambda(x)-\frac1n\right\}
\]
is analytic and has nonempty sections.  The Jankov--von Neumann selection theorem yields a
universally measurable map $T_n$ with $(x,T_n(x))\in A_n$.  Therefore
\[
 f(T_n(x))-\lambda c(x,T_n(x))
 \ge \varphi_\lambda(x)-\frac1n.
\]
The coupling $(\Id,T_n)_\#\mu$ realizes the pointwise envelope up to $1/n$.  Truncation and Fatou's
lemma handle the possible lack of a common integrable lower bound.  This selection argument is the
mechanism by which pointwise robust responses are converted into admissible transport plans.

\section{Dual attainment}
Define the one-dimensional objective
\[
 g(\lambda):=\lambda\delta+\int\varphi_\lambda\,d\mu.
\]
It is convex as a supremum of affine functions followed by integration.  It is lower semicontinuous:
if $\lambda_n\to\lambda$, upper semicontinuity of $f-\lambda c$ and the measurable-envelope
construction yield the required liminf inequality.  Also
\[
 \varphi_\lambda(x)\ge f(x)
\]
because $c(x,x)=0$, so
\[
 g(\lambda)\ge \lambda\delta+\int f\,d\mu\longrightarrow+\infty
 \qquad(\lambda\to\infty).
\]
Thus $g$ attains its minimum at some finite $\lambda^*\ge0$, and
$(\lambda^*,\varphi_{\lambda^*})$ is a dual optimizer.

\section{Complementary slackness}
For any feasible $\pi$ and dual optimizer,
\begin{align*}
 0&\le \int\bigl(\varphi_{\lambda^*}(x)+\lambda^*c(x,y)-f(y)\bigr)\,d\pi,\\
 0&\le \lambda^*\left(\delta-\int c\,d\pi\right).
\end{align*}
The sum of these two nonnegative quantities is exactly the dual objective minus the primal
objective.  Equality of the primal and dual values therefore holds if and only if each term vanishes.
The first integrand is nonnegative, so its integral is zero exactly when the contact condition
\eqref{eq:BM-contact} holds $\pi$-almost surely.  The second term is precisely
\eqref{eq:BM-slack}.  This proves the optimizer characterization.

\section{Interpretation}
At a positive optimal multiplier, an optimal coupling spends the full transport budget and is
concentrated on pairs $(x,y)$ for which $y$ maximizes the penalized payoff
$f(y)-\lambda^*c(x,y)$.  If the optimal multiplier is zero, the budget may be slack and the coupling
is supported on global maximizers of $f$.  Thus the theorem is both a scalar duality formula and a
structural description of worst-case transport.

\input{references_manual}
\end{document}
