
\input{preamble}
\title{Step-by-step modernization of Carroll (2004): Birkhoff's contraction coefficient}
\author{}
\date{Source PDF: \texttt{BirkhoffHilbert-Carroll2005-carroll-birkoff.pdf}}
\begin{document}
\maketitle

\section*{Source map and scope}
\begin{description}[style=nextline,leftmargin=3.2cm,labelwidth=3cm]
\item[Primary source] \texttt{\detokenize{BirkhoffHilbert-Carroll2005-carroll-birkoff.pdf}}
\item[Coverage] Theorem 1 and the complete finite-dimensional linear-programming reduction proving the sharp Birkhoff contraction coefficient.
\item[Supplemental references] Carroll explicitly cites Luenberger for the basic-feasible-solution theorem used in Section 2.2 \cite{Luenberger1984}.  No later contraction proof is substituted.
\item[Source anchors]
\begin{itemize}[leftmargin=2em]
  \item Theorem 1
  \item Sections 2.1--2.4
  \item Remark 2
\end{itemize}
\end{description}

The exposition below preserves the source proof order and proof technique.  Notation is normalized
and intermediate calculations are made explicit.  When the source contains a gap, ambiguity, or
apparently incorrect step, the reconstruction records the source step before adding a separate
commentary remark.

\section*{Reference and role}
This note modernizes Joseph E. Carroll, \emph{Birkhoff's Contraction Coefficient},
Linear Algebra and its Applications 389:227--234, 2004 \cite{Carroll2004}.  Carroll
proves Birkhoff's finite positive-matrix contraction theorem by elementary algebra,
calculus, and a small linear-programming reduction.  This is a high-value companion
to Eveson--Nussbaum because it avoids general cone topology and gives a finite
coordinate proof of the exact contraction coefficient.

The paper uses a multiplicative version of Hilbert's projective distance.  Taking
logarithms converts it into the more common additive Hilbert metric and turns
multiplicative powers into ordinary contraction inequalities.

\section*{Definitions in modern notation}
For $n\ge2$, let
\[
  \R^n_{++}:=\{x\in\R^n:x_i>0\text{ for all }i\}.
\]
This is an abelian group under coordinatewise multiplication.  Carroll defines the
multiplicative projective norm
\[
  \|x\|:=\frac{\max_i x_i}{\min_i x_i}
        =\max_{i,j}\frac{x_i}{x_j}
  \qquad (x\in\R^n_{++}).
\]
It satisfies:
\begin{enumerate}[leftmargin=2em]
\item $\|x\|=1$ iff $x$ is a constant positive vector;
\item $\|x^{-1}\|=\|x\|$;
\item $\|xy\|\le \|x\|\|y\|$;
\item $\|cx\|=\|x\|$ for every scalar $c>0$.
\end{enumerate}
The multiplicative projective distance is
\[
  d(x,y):=\|xy^{-1}\|.
\]
Thus
\[
  d(x,y)=\frac{\max_i(x_i/y_i)}{\min_i(x_i/y_i)}.
\]
It satisfies the multiplicative triangle inequality
\[
  d(x,z)\le d(x,y)d(y,z),
\]
and it is projective:
\[
  d(ax,by)=d(x,y)\qquad(a,b>0).
\]
The additive Hilbert projective pseudometric is
\[
  \delta(x,y):=\log d(x,y).
\]

For $r\ge1$, Carroll writes
\[
  \tau(r):=\frac{\sqrt r-1}{\sqrt r+1}.
\]
The map $\tau:[1,\infty)\to[0,1)$ is strictly increasing and onto.

Let $A$ be a positive $m\times n$ matrix, with rows $a_1,\dots,a_m\in\R^n_{++}$.
Carroll defines the multiplicative diameter of the row set by
\[
  d(A):=\max_{1\le i,j\le m}d(a_i,a_j).
\]
Written in coordinates,
\[
  d(A)=\max_{i,j,k,\ell}\frac{a_{ik}a_{j\ell}}{a_{i\ell}a_{jk}}.
\]
This is the same multiplicative cross-ratio diameter that appears in finite versions
of Birkhoff--Hopf.  Its logarithm is the additive Hilbert diameter of the image.

\section*{Theorem 1 in additive and multiplicative forms}
Carroll's Theorem 1 says:

\begin{quote}
Let $m,n\ge2$, let $A$ be a positive $m\times n$ matrix, and assume $d(A)>1$.
For all $x,y\in\R^n_{++}$ with $d(x,y)>1$,
\[
  d(Ax,Ay)<d(x,y)^{\tau(d(A))}.
\]
Moreover, if $\eta<\tau(d(A))$, then there exist $x,y\in\R^n_{++}$ such that
\[
  d(Ax,Ay)>d(x,y)^\eta.
\]
\end{quote}

Taking logarithms gives the usual Hilbert-metric statement:
\[
  \delta(Ax,Ay)<\tau(d(A))\,\delta(x,y),
\]
and the second assertion says that the coefficient $\tau(d(A))$ is minimal.  If
$D=\log d(A)$ is the additive diameter, then
\[
  \tau(d(A))=\frac{e^{D/2}-1}{e^{D/2}+1}=\tanh(D/4).
\]

The proof has four stages:
\begin{enumerate}[leftmargin=2em]
\item reduce the number of output rows to $2$;
\item rewrite the $2\times n$ problem as an inequality for a scalar function $F$;
\item reduce the $F$-inequality to the $2\times2$ case by linear programming;
\item solve the $2\times2$ inequality and prove minimality.
\end{enumerate}

\sourceanchor{Section 2.1, reduction to the $2\times n$ case and inequality (1a).}
\section*{Step 1: reduction to two output rows}
Assume the theorem is known for matrices with two rows.  Let $A$ be an arbitrary
positive $m\times n$ matrix and take $x,y\in\R^n_{++}$ with $d(Ax,Ay)>1$.
By the definition of $d(Ax,Ay)$, there are row indices $i,j$ such that
\[
  d(Ax,Ay)
  =\frac{(a_i\cdot x)(a_i\cdot y)^{-1}}
         {(a_j\cdot x)(a_j\cdot y)^{-1}}
  = \frac{(a_i\cdot x)(a_j\cdot y)}{(a_i\cdot y)(a_j\cdot x)},
\]
after choosing the order so the displayed ratio is the maximum and is $>1$.
Let $A'$ be the $2\times n$ submatrix whose rows are $a_i$ and $a_j$.  Then
\[
  d(A'x,A'y)=d(Ax,Ay)>1.
\]
Also
\[
  d(A')=d(a_i,a_j)\le d(A),
\]
and $\tau$ is increasing.  Applying the two-row theorem to $A'$ gives
\[
  d(Ax,Ay)=d(A'x,A'y)
  <d(x,y)^{\tau(d(A'))}
  \le d(x,y)^{\tau(d(A))}.
\]
So it remains to prove the theorem for $m=2$.

\sourceanchor{Section 2.1, restatement as inequality (1b).}
\section*{Step 2: the two-row problem and the function $F$}
Let the two rows of $A$ be $a,b\in\R^n_{++}$.  We must prove
\[
  d(Ax,Ay)<d(x,y)^{\tau(d(a,b))}.
\]
The distance $d(Ax,Ay)$ is the larger of the ratio
\[
  \frac{(a\cdot x)(a\cdot y)^{-1}}{(b\cdot x)(b\cdot y)^{-1}}
\]
and its reciprocal.  Swapping the labels $a$ and $b$ takes the reciprocal and leaves
$d(a,b)$ unchanged.  Therefore it is enough to prove
\[
  \frac{(a\cdot x)(a\cdot y)^{-1}}{(b\cdot x)(b\cdot y)^{-1}}
  <d(x,y)^{\tau(d(a,b))}
\]
for all $a,b,x,y$ with $d(a,b)>1$ and $d(x,y)>1$.

Now use coordinatewise multiplication by $b^{-1}$.  This is an isometry for $d$, so
$d(a,b)=\|ab^{-1}\|$.  Define
\[
  r:=ab^{-1},
  \qquad
  s:=xy^{-1}.
\]
Then $r$ and $s$ are nonconstant, $\|r\|=d(a,b)$, and $\|s\|=d(x,y)$.  Since
$x=sy$,
\begin{align*}
  \frac{(a\cdot x)(a\cdot y)^{-1}}{(b\cdot x)(b\cdot y)^{-1}}
  &=\frac{(ab^{-1}\cdot x)(ab^{-1}\cdot y)^{-1}}
          {(1\cdot x)(1\cdot y)^{-1}}\\
  &=\frac{(r\cdot sy)(1\cdot y)}{(r\cdot y)(1\cdot sy)}\\
  &=\frac{(rs\cdot y)(1\cdot y)}{(r\cdot y)(s\cdot y)}.
\end{align*}
This motivates Carroll's function
\[
  F(r,s,y):=\frac{(rs\cdot y)(1\cdot y)}{(r\cdot y)(s\cdot y)}.
\]
The desired two-row inequality is therefore equivalent to
\[
  F(r,s,y)<\|s\|^{\tau(\|r\|)}
\]
for all nonconstant $r,s\in\R^n_{++}$ and all $y\in\R^n_{++}$.  If $F(r,s,y)\le1$,
there is nothing to prove, since $\|s\|^{\tau(\|r\|)}>1$.  Hence Carroll restricts to
$F(r,s,y)>1$.

\sourceanchor{Section 2.2, the linear program and the basic-feasible-solution reduction.}
\section*{Step 3: properties of $F$ and reduction to two coordinates}
Carroll broadens the $y$-domain temporarily from $\R^n_{++}$ to the closed cone
$\R^n_+$ so that a linear-programming maximum may occur on the boundary.

The function $F$ is homogeneous of degree zero in each variable:
\[
  F(\alpha r,\beta s,\gamma y)=F(r,s,y)
  \qquad(\alpha,\beta,\gamma>0).
\]
The right-hand side $\|s\|^{\tau(\|r\|)}$ is also unchanged by positive rescaling of
$r$ or $s$.  Therefore we may rescale $r$ and $s$ so that
\[
  r\cdot y=1,
  \qquad
  s\cdot y=1\cdot y.
\]
The second condition is equivalent to
\[
  (s-1)\cdot y=0.
\]
Under these two normalizations,
\[
  F(r,s,y)=rs\cdot y.
\]
Indeed,
\[
  F(r,s,y)=\frac{(rs\cdot y)(1\cdot y)}{1\cdot(1\cdot y)}=rs\cdot y.
\]

Now consider the linear program
\[
  \text{maximize } rs\cdot x
\]
subject to
\[
  x\in\R^n_+,
  \qquad
  r\cdot x=1,
  \qquad
  (s-1)\cdot x=0.
\]
The vector $y$ is feasible.  The feasible set is bounded because $r_i>0$ and
$r\cdot x=1$ imply $0\le x_i\le 1/r_i$.  Hence a maximum exists.  By the fundamental
basic-feasible-solution theorem of linear programming, a maximum is attained at a
basic feasible solution $x^0$ with at most two nonzero coordinates, because there are
two equality constraints.

If $x^0$ had only one nonzero coordinate, then the corresponding one-dimensional
version of $F$ would equal $1$, contradicting the case $F(r,s,y)>1$ after taking a
maximizer that does at least as well as $y$.  Thus there are distinct indices $i,j$ and
positive numbers $x_i,x_j$ such that
\[
  F(r,s,y)\le F((r_i,r_j),(s_i,s_j),(x_i,x_j)).
\]
Moreover,
\[
  \|(r_i,r_j)\|\le \|r\|,
  \qquad
  \|(s_i,s_j)\|\le \|s\|.
\]
Since $\tau$ is increasing, a proof of the desired inequality in dimension $2$ implies
the desired inequality in dimension $n$.

This is the key finite-dimensional reduction: the worst case lives on two coordinates.

\sourceanchor{Section 2.3, the $2\times2$ calculation.}
\section*{Step 4: the two-coordinate calculation}
Let $r,s\in\R^2_{++}$ be nonconstant and $y\in\R^2_{++}$.  By permuting coordinates
and rescaling each variable, assume
\[
  r=(1,r),\quad r>1,
  \qquad
  s=(1,s),\quad s\ne1,
  \qquad
  y=(1,y),\quad y>0.
\]
Assume again that $F(r,s,y)>1$.  Then
\begin{align*}
  F(r,s,y)
  &=\frac{((1,rs)\cdot(1,y))((1,1)\cdot(1,y))}
          {((1,r)\cdot(1,y))((1,s)\cdot(1,y))}\\
  &=\frac{(1+rsy)(1+y)}{(1+ry)(1+sy)}\\
  &=\frac{1+rsy^2+(1+rs)y}{1+rsy^2+(r+s)y}.
\end{align*}
If $s<1$, then
\[
  (1+rs)-(r+s)=(r-1)(s-1)<0,
\]
so the numerator is smaller than the denominator and $F(r,s,y)<1$, contrary to the
standing assumption.  Hence $s>1$, and $\|s\|=s$.

Use the square inequality
\[
  0\le (1-y\sqrt{rs})^2=1+rsy^2-2y\sqrt{rs}.
\]
Thus
\[
  1+rsy^2\ge 2y\sqrt{rs}.
\]
Since $1+rs<r+s$ is false when $r,s>1$; rather $1+rs-(r+s)=(r-1)(s-1)>0$, increasing
the common term $1+rsy^2$ in the fraction
\[
  \frac{1+rsy^2+(1+rs)y}{1+rsy^2+(r+s)y}
\]
would decrease the fraction.  Therefore replacing $1+rsy^2$ by its lower bound gives
an upper bound:
\begin{align*}
  F(r,s,y)
  &\le
  \frac{2y\sqrt{rs}+(1+rs)y}{2y\sqrt{rs}+(r+s)y}\\
  &=\frac{2\sqrt{rs}+1+rs}{2\sqrt{rs}+r+s}\\
  &=\left(\frac{1+\sqrt{rs}}{\sqrt r+\sqrt s}\right)^2.
\end{align*}
Equality occurs when $y\sqrt{rs}=1$.

It remains to prove
\[
  \left(\frac{1+\sqrt{rs}}{\sqrt r+\sqrt s}\right)^2
  < s^{\tau(r)}.
\]
Taking square roots and renaming $\sqrt r$ and $\sqrt s$ as $r$ and $s$, it is enough
to prove, for all $r,s>1$,
\[
  f_1(r,s):=\frac{1+rs}{r+s}
  < s^{(r-1)/(r+1)}=:g_1(r,s).
\]
At $s=1$ both sides equal $1$.  By the mean value theorem it suffices to show
\[
  \frac{\partial f_1}{\partial s}(r,s)
  <
  \frac{\partial g_1}{\partial s}(r,s)
  \qquad(s>1).
\]
Compute
\[
  \frac{\partial f_1}{\partial s}
  =\frac{r^2-1}{(r+s)^2},
\]
and
\[
  \frac{\partial g_1}{\partial s}
  =\frac{r-1}{r+1}s^{-2/(r+1)}.
\]
Thus the desired derivative inequality is
\[
  \frac{r^2-1}{(r+s)^2}
  <\frac{r-1}{r+1}s^{-2/(r+1)}.
\]
Canceling the positive factor $(r-1)/(r+1)$, this is equivalent to
\[
  \left(\frac{r+1}{r+s}\right)^2<s^{-2/(r+1)}.
\]
Taking positive square roots and rearranging, it is enough to show
\[
  f_2(r,s):=\frac{r+s}{r+1}>s^{1/(r+1)}=:g_2(r,s).
\]
Again $f_2(r,1)=g_2(r,1)=1$.  And
\[
  \frac{\partial f_2}{\partial s}=\frac{1}{r+1},
  \qquad
  \frac{\partial g_2}{\partial s}=\frac{1}{r+1}s^{-r/(r+1)}.
\]
For $s>1$, $s^{-r/(r+1)}<1$, hence
\[
  \frac{\partial f_2}{\partial s}>
  \frac{\partial g_2}{\partial s}.
\]
Therefore $f_2>g_2$ for $s>1$, so the derivative inequality for $f_1,g_1$ holds,
so $f_1<g_1$, and consequently
\[
  F(r,s,y)<\|s\|^{\tau(\|r\|)}.
\]
This proves that $\tau(d(A))$ is a contraction coefficient.

\sourceanchor{Section 2.4, minimality of $\tau(d(A))$.}
\section*{Step 5: minimality of the coefficient}
Carroll then proves that no smaller coefficient works.  The proof reverses the
sharpness points in the two-coordinate calculation.

First consider the $2\times2$ $F$-problem.  Let $r=(r_1,r_2)$ be nonconstant.  After
rescaling and switching coordinates, assume
\[
  r=(1,R),\qquad R>1.
\]
Let $0<\eta<\tau(R)$.  Since $\tau$ is strictly increasing and onto $[0,1)$, choose
$R'$ with
\[
  1<R'<R,
  \qquad
  \eta=\tau(R').
\]
In the proof above, equality in the AM--GM step is obtained by taking
\[
  y=(1,(Rs)^{-1/2})
\]
when $s=(1,s)$ with $s>1$.  The square-root inequality then compares
\[
  f_1(\sqrt R,\sqrt s)
  \quad\text{with}\quad
  (\sqrt s)^\eta.
\]
At $s=1$ the two sides agree, but the derivative of the left side is larger than the
derivative corresponding to the smaller exponent $\eta=\tau(R')$.  Hence for all
$s>1$ sufficiently close to $1$,
\[
  F\left((1,R),(1,s),(1,(Rs)^{-1/2})\right)>s^\eta.
\]
Since $\|(1,s)\|=s$, this says there are positive $s,y$ such that
\[
  F(r,s,y)>\|s\|^\eta.
\]
Thus any exponent smaller than $\tau(\|r\|)$ fails already in dimension $2$.

For $n\ge3$, choose two coordinates where $r$ attains its minimum and maximum.
After scaling and permuting, assume
\[
  r_1=1,
  \qquad
  r_2=\|r\|=R>1.
\]
Use the two-dimensional construction on the first two coordinates.  Initially put the
remaining components of $y$ equal to $0$ so that the value of $F$ is exactly the
$2$-coordinate value.  Then perturb those zero components to sufficiently small
positive values.  The function $F$ is continuous in $y$ on the region where the
denominator is nonzero, so the strict inequality
\[
  F(r,s,y)>\|s\|^\eta
\]
persists for a positive $y\in\R^n_{++}$.  This proves minimality for the $F$-problem
in every dimension.

Finally, for an arbitrary positive $m\times n$ matrix $A$, choose rows $a_i,a_j$ with
\[
  d(a_i,a_j)=d(A).
\]
The two-row minimality construction applied to these rows gives vectors $x,y$ for
which the corresponding two-row image violates any exponent $\eta<\tau(d(A))$.
The full matrix image has distance at least as large as the distance seen by this row
pair, so the same $x,y$ prove minimality for $A$.

\section*{The two closing remarks of the paper}
\sourceanchor{Remarks 2 and 3.}

\paragraph{Comparable vectors with zeros.}
Carroll observes that the same proof extends from strictly positive vectors to vectors in $\mathbb{R}_+^n$ having at least two positive coordinates, provided the two vectors have zeros in exactly the same positions.  The multiplicative spread is computed only over their common positive support.  Every reduction in Sections 2.1--2.4 stays inside that support, so the argument is unchanged.

\paragraph{The symmetric bonus inequality.}
Because $F(r,s,y)=F(s,r,y)$, the proof also gives
\[
 F(r,s,y)<\|r\|^{\tau(\|s\|)}.
\]
Translating this back to matrices yields
\[
 d(Ax,Ay)<d(A)^{\tau(d(x,y))}.
\]
Carroll notes no direct use for this sharper symmetric companion.  It implies $d(Ax,Ay)<d(A)$, hence boundedness of the positive image in Hilbert projective distance, but that weaker fact has a direct proof.

\section*{Translation to the finite Birkhoff--Hopf formula}
Let
\[
  \Theta(A):=d(A)=\max_{i,j,k,\ell}\frac{a_{ik}a_{j\ell}}{a_{i\ell}a_{jk}}.
\]
Then Carroll proves, in additive Hilbert metric,
\[
  \delta(Ax,Ay)
  \le \frac{\sqrt{\Theta(A)}-1}{\sqrt{\Theta(A)}+1}\,\delta(x,y),
\]
with sharp coefficient.  If
\[
  \Delta(A):=\log\Theta(A),
\]
then
\[
  \frac{\sqrt{\Theta(A)}-1}{\sqrt{\Theta(A)}+1}
  =\frac{e^{\Delta(A)/2}-1}{e^{\Delta(A)/2}+1}
  =\tanh(\Delta(A)/4).
\]
This is exactly the finite positive-matrix Birkhoff contraction coefficient used by
Franklin--Lorenz, just written in Carroll's multiplicative notation.

\begin{remark}[Strict and non-strict formulations]
Carroll proves a strict inequality whenever both input and matrix diameters exceed one.  The global Lipschitz coefficient is conventionally written with $\le$ because it is the supremum over all pairs; minimality shows that this supremum equals $\tau(d(A))$, although no fixed nondegenerate pair need attain equality.
\end{remark}

\section*{Proof dependency chain}
Carroll's proof reduces the theorem to the following sequence.
\begin{commentarynotes}
\item Define the multiplicative spread of a positive finite vector:
\[
  \|x\|=\max_{i,j}x_i/x_j.
\]
\item Define the multiplicative Hilbert distance
\[
  d(x,y)=\|(x_i/y_i)_i\|.
\]
The additive version is its logarithm.
\item Define the matrix cross-ratio diameter
\[
  \Theta(A)=\max_{i,j,k,\ell}\frac{a_{ik}a_{j\ell}}{a_{i\ell}a_{jk}}.
\]
\item Prove that positive diagonal row and column scalings preserve $\Theta(A)$.
\item Prove the finite contraction theorem
\[
  \log d(Ax,Ay)
  \le \tau(\Theta(A))\log d(x,y).
\]
For applications requiring only a strict contraction, the exact sharpness part can be separated; existence of some
$\rho<1$ follows from strict positivity and finiteness.
\item Use Franklin--Lorenz Lemma 2 to apply this theorem twice per Sinkhorn cycle,
producing the squared factor $\gamma=\tau(\Theta(A))^2$.
\end{commentarynotes}

\paragraph{Why Carroll is high value.}
Eveson--Nussbaum gives a very general cone-theoretic theorem.  Carroll gives a
coordinate finite-dimensional proof, including the exact coefficient, with only one
linear-programming reduction and elementary two-variable inequalities.  This makes it
a direct finite-dimensional source for the sharp theorem specialized to strictly positive matrices.

\input{references_manual}
\end{document}
