\input{preamble}
\title{Modernized Proof Notes for Fortet (1940): Resolution of Schrodinger's System}
\author{}
\date{Source PDF: \texttt{Fortet 1940 - JMPA\_1940\_9\_19\_1-4\_83\_0.pdf}}
\begin{document}
\maketitle

\section*{Source map and scope}
\begin{description}[style=nextline,leftmargin=3.2cm,labelwidth=3cm]
\item[Primary source] \texttt{\detokenize{Fortet 1940 - JMPA_1940_9_19_1-4_83_0.pdf}}
\item[Coverage] Fortet's existence and uniqueness proof for the Schrödinger system, including truncation, monotone iteration, mass identities, limiting arguments, and Bernstein's conclusion.
\item[Source anchors]
\begin{itemize}[leftmargin=2em]
  \item Theorems I--III
  \item successive approximation construction
  \item lemmas controlling truncation and limits
  \item final Bernstein application
\end{itemize}
\end{description}

The exposition below preserves the source proof order and proof technique.  Notation is normalized
and intermediate calculations are made explicit.  When the source contains a gap, ambiguity, or
apparently incorrect step, the reconstruction records the source step before adding a separate
commentary remark.

\section*{Reference and purpose}
This note expands Robert Fortet, ``Resolution d'un systeme d'equations de M. Schrodinger,'' \emph{Journal de mathematiques pures et appliquees}, 9e serie, tome 19, pp. 83--105, 1940 \cite{Fortet1940}.
The original is written in older French analytic language.  The reconstruction keeps the proof architecture, translates the symbols into modern measure-theoretic notation, and records the almost-everywhere conventions explicitly.

\paragraph{Modernization convention.}
Fortet writes two real intervals as $J'$ and $J''$ and writes the marginal densities as $\omega_1(x)$ and $\omega_2(y)$.  We keep this notation.  Integrals are Lebesgue integrals; ``measurable (B)'' means Borel measurable.  Equalities between functions are to be read outside null sets when the original proof says ``presque partout.''

\section{The Schrodinger system}
Fortet studies a nonnegative kernel $g(x,y)$ on $J'\times J''$ and two nonnegative marginal densities $\omega_1,\omega_2$.  His system is
\begin{align}
  \varphi(x) \int_{J''} g(x,y)\psi(y)\,dy &= \omega_1(x), \tag{S1}\label{eq:fortet-S1}\\
  \psi(y) \int_{J'} g(x,y)\varphi(x)\,dx &= \omega_2(y). \tag{S2}\label{eq:fortet-S2}
\end{align}
A solution is considered only up to the gauge
\[
  (\varphi,\psi) \sim (c\varphi,c^{-1}\psi), \qquad c\ne 0.
\]
Fortet calls a solution positive when $\varphi\ge0$ and $\psi\ge0$; he is mainly interested in such solutions.

\subsection*{Hypotheses I and II, as used by Fortet}
Fortet separates the data assumptions into two groups.  His Hypotheses I are
\[
  \omega_1(x)\ge 0,\qquad \omega_2(y)\ge 0,
\]
with both masses finite, positive, and equal:
\[
  0<\int_{J'}\omega_1(x)\,dx
    =\int_{J''}\omega_2(y)\,dy<\infty.
\]
He then normalizes the common mass to one:
\begin{equation}
  \int_{J'}\omega_1(x)\,dx
  =\int_{J''}\omega_2(y)\,dy=1. \tag{3'}\label{eq:fortet-normalization}
\end{equation}

His standing Hypotheses II are:
\begin{enumerate}[leftmargin=2em]
\item $g:J'\times J''\to[0,\infty)$ is continuous and bounded above by a
finite positive constant $\lambda$;
\item for every fixed $x$, the function $y\mapsto g(x,y)$ is nonzero except
possibly on a null set, and for every fixed $y$, the function
$x\mapsto g(x,y)$ is nonzero except possibly on a null set;
\item $\omega_1$ and $\omega_2$ are continuous.
\end{enumerate}
These are the assumptions used in the existence and uniqueness sections.  The
paper later remarks that continuity of the marginals can be weakened to Borel
measurability with small changes.

\begin{remark}[Relation to later reconstructions]
L\'eonard's reconstruction \cite{Leonard2019} and the closely source-following
account of Essid--Pavon make explicit several measurability and monotone-limit
steps that Fortet compresses.  Whenever such a later argument is used below, it
is labeled as supplemental; the iteration and comparison mechanism remains
Fortet's own.
\end{remark}

\section{Reduction to one scalar fixed-point equation}
From \eqref{eq:fortet-S2}, whenever the denominator is meaningful,
\[
  \psi(y)=\frac{\omega_2(y)}{\int_{J'}g(z,y)\varphi(z)\,dz}.
\]
Substituting this into \eqref{eq:fortet-S1} gives a single nonlinear equation for $\varphi$.
Fortet then changes variables by setting
\[
  h(x)=\frac{\omega_1(x)}{\varphi(x)}.
\]
Then $\varphi(x)=\omega_1(x)/h(x)$ and the single equation becomes
\begin{equation}
  h(x)=Q(h)(x)
  :=\int_{J''} g(x,y)
    \frac{\omega_2(y)}{G(h,y)}\,dy, \tag{F}\label{eq:fortet-fixed-point}
\end{equation}
where
\begin{equation}
  G(h,y):=\int_{J'}g(z,y)\frac{\omega_1(z)}{h(z)}\,dz. \tag{G}\label{eq:fortet-G}
\end{equation}
A nonzero, noninfinite solution of \eqref{eq:fortet-fixed-point} gives a solution of the original system by
\begin{equation}
  \varphi(x)=\frac{\omega_1(x)}{h(x)},
  \qquad
  \psi(y)=\frac{\omega_2(y)}{G(h,y)}. \tag{Recover}\label{eq:fortet-recover}
\end{equation}
This is the exact continuum ancestor of the alternating quotient updates in Sinkhorn scaling: divide one marginal by the current kernel-propagated scaling, then divide the other marginal by the propagated result.

\section{Fortet's class $(C)$}
Fortet introduces a class $(C)$ of functions on $J'$.
In modern notation, $H\in(C)$ if:
\begin{enumerate}[leftmargin=2em]
\item $H$ is Borel measurable;
\item $H$ is bounded below by a strictly positive constant;
\item $H$ is finite almost everywhere.
\end{enumerate}
The lower bound prevents the denominator in \eqref{eq:fortet-G} from blowing up by division by $H$.
The finiteness almost everywhere is enough because all equations are ultimately read modulo null sets on the supports of the marginals.

\section{The fundamental lemma}
Let
\[
  A_1:=\{x\in J': \omega_1(x)>0\}.
\]
For $H\in(C)$ Fortet forms $H'=Q(H)$ by \eqref{eq:fortet-fixed-point}.  The central lemma is the engine of the paper.

\begin{lemma}[Fortet's fixed-point comparison lemma, modern form]
If $H\in(C)$ and $H'=Q(H)$, then $H'$ is Borel measurable, is bounded below by a positive number on every finite subinterval of $J'$, and is finite almost everywhere on $A_1$.  Moreover, if either
\[
  H'\le H \quad\hbox{a.e.}
  \qquad\hbox{or}\qquad
  H'\ge H \quad\hbox{a.e.},
\]
then
\[
  H'=H \quad\hbox{a.e. on } A_1.
\]
\end{lemma}

\subsection{Proof of the regularity part}
Fortet uses finite truncations.  Choose increasing finite sets or finite intervals
\[
  J'_1\subset J'_2\subset\cdots\uparrow J',
  \qquad
  J''_1\subset J''_2\subset\cdots\uparrow J''.
\]
For $H\in(C)$ define truncated denominators
\[
  G_n(H,y)=\int_{J'_n} g(z,y)\frac{\omega_1(z)}{H(z)}\,dz.
\]
Because $g$ and $\omega_1$ are measurable and $H$ is positive and finite a.e., the integrand is nonnegative and measurable.  The functions $G_n(H,y)$ are well-defined and increase in $n$.
Let
\[
  G(H,y)=\lim_{n\to\infty}G_n(H,y).
\]
Fortet records that $G(H,\cdot)$ is of Baire class one, hence Borel measurable, positive on finite pieces in the required sense, and finite where the hypotheses force it to be finite.  The finite-piece lower bound comes from the nontriviality of $g$ on each finite interval together with the positive lower bound on $H$.

Then define truncated images
\[
  H'_n(x)=\int_{J''_n} g(x,y)\frac{\omega_2(y)}{G(H,y)}\,dy.
\]
The monotone convergence theorem gives
\[
  H'(x)=\lim_{n\to\infty}H'_n(x).
\]
Thus $H'$ is measurable.  Positivity and local lower bounds are inherited
from the positivity and nontriviality of the kernel and marginal on finite
intervals.  Finiteness almost everywhere on $A_1$ is not asserted directly
from pointwise estimates; it is extracted from the mass identity below.

\subsection{The normalization identity}
Fortet defines the auxiliary nonnegative function
\[
  F(x,y)=g(x,y)\frac{\omega_2(y)}{G(H,y)}\frac{\omega_1(x)}{H(x)}.
\]
Truncating in both variables and applying Beppo-Levi/monotone convergence gives
\begin{align*}
  \int_{J'} \frac{H'(x)}{H(x)}\omega_1(x)\,dx
  &= \int_{J'}\left(\int_{J''} g(x,y)\frac{\omega_2(y)}{G(H,y)}\,dy\right)
       \frac{\omega_1(x)}{H(x)}\,dx \\
  &= \int_{J''}\frac{\omega_2(y)}{G(H,y)}
       \left(\int_{J'}g(x,y)\frac{\omega_1(x)}{H(x)}\,dx\right)dy \\
  &= \int_{J''}\omega_2(y)\,dy \\
  &=1. \tag{M}\label{eq:fortet-mass-identity}
\end{align*}
This identity is Fortet's equation (12) in modern notation.  It implies $H'$ is finite a.e. on $A_1$, because if $H'=\infty$ on a set of positive $\omega_1$-mass, the integral in \eqref{eq:fortet-mass-identity} would be infinite.

\subsection{The comparison conclusion}
If $H'\le H$ a.e., then
\[
  0\le \frac{H'}{H}\le 1 \quad\hbox{a.e.}
\]
and \eqref{eq:fortet-mass-identity} plus $\int\omega_1=1$ forces $H'/H=1$ $\omega_1$-a.e.  Therefore $H'=H$ on $A_1$ up to a null set.
Similarly, if $H'\ge H$ a.e., then $H'/H\ge1$ and the same integral identity forces equality.
This is the central fixed-point forcing step: one-sided comparison plus the mass identity collapses to equality.

\section{Successive approximation: the first existence theorem}

\subsection{The exact clipped iteration on source page 90}
Fortet writes $M(f,g)=\max\{f,g\}$ and $m(f,g)=\min\{f,g\}$ pointwise.
Starting from the constant function $H_1=1$, he defines three sequences:
\begin{align*}
  H_1'&=Q(H_1),
  &H_1''&=m(H_1,H_1'),\\
  H_n&=M\!\left(H_{n-1}'',\frac1n\right) \quad (n\ge2),
  &H_n'&=Q(H_n),
  &H_n''&=m(H_n,H_n').
\end{align*}
Thus the first two rounds are literally
\[
  H_1=1,\quad H_1'=Q(H_1),\quad H_1''=m(H_1,H_1'),
\]
\[
  H_2=M(H_1'',1/2),\quad H_2'=Q(H_2),\quad
  H_2''=m(H_2,H_2').
\]
The floor $1/n$ keeps every input $H_n$ in class $(C)$, while the minimum with
$Q(H_n)$ creates the monotone envelope.  Fortet checks directly that
\begin{equation}
  H_{n+1}\le H_n,
  \qquad
  H_{n+1}''\le H_n''. \tag{I}\label{eq:fortet-iterate-monotone}
\end{equation}
The verification is inductive.  Since $H_n''\le H_n$ and
$1/(n+1)\le1/n$, one has
\[
  H_{n+1}=M(H_n'',1/(n+1))\le M(H_n,1/n)=H_n.
\]
The map $Q$ is order preserving: $H\le K$ implies
$G(H,\cdot)\ge G(K,\cdot)$ and hence $Q(H)\le Q(K)$.  Therefore
$H_{n+1}'\le H_n'$, and taking minima gives the second inequality in
\eqref{eq:fortet-iterate-monotone}.

\begin{remark}[Modern compact notation]
The later formulation
\[
  u_{n+1}=\frac{U}{n+1}\vee\bigl(\Phi[u_n]\wedge U\bigr)
\]
used by L\'eonard \cite[Eq.~(3.3)]{Leonard2019} packages the same clipping
idea into a single recurrence.  Fortet's paper uses the three displayed
sequences $H_n,H_n',H_n''$ above, and that is the notation retained in the
proof reconstruction.
\end{remark}

\subsection{First case: an ordered iterate occurs}
Assume that for some $n_0$,
\[
  H_{n_0}'\le H_{n_0}
  \quad\text{almost everywhere}.
\]
Then $H_{n_0}''=H_{n_0}'$ almost everywhere.  The function
$H_{n_0}'$ may fail the strict positive lower bound required in class $(C)$,
so Fortet regularizes it by
\[
  K_p=M\!\left(H_{n_0}',\frac1p\right),
  \qquad K_p'=Q(K_p).
\]
As $p\to\infty$, $K_p\downarrow H_{n_0}'$.  Because $Q$ is increasing,
$K_p'$ decreases to a measurable limit $K'$, and Beppo--Levi applied twice
gives
\[
  K'=\lim_{p\to\infty}Q(K_p)=Q(H_{n_0}')
\]

after interpreting the reciprocal denominators with the paper's nonnegative
integral conventions.

The monotone bounds used by Fortet imply
\[
  H_{n_0}'\le K_p\le H_{n_0},
  \qquad
  H_{n_0}'\le K_p'\le H_{n_0+1}'
\]
together with a uniform integrability estimate for the quotients $K_p'/K_p$.
For each $p$, the mass identity gives
\[
  \int_{J'}\frac{K_p'(x)}{K_p(x)}\omega_1(x)\,dx=1.
\]
Passing to the limit yields Fortet's equation (14):
\begin{equation}
  \int_{J'}\frac{K'(x)}{H_{n_0}'(x)}\omega_1(x)\,dx=1. \tag{14}
\end{equation}
The order inequalities give $K'\le H_{n_0}'$.  Hence the quotient in
(14) is at most one, and because $\int\omega_1=1$ it must equal one
almost everywhere on $A_1=\{\omega_1>0\}$.  Thus
\[
  K'=H_{n_0}' \quad\text{a.e. on }A_1.
\]
The defining denominator calculation then shows
\[
  Q(K')=Q(H_{n_0}')=K',
\]
so $K'$ is a positive fixed point of $Q$ and
\eqref{eq:fortet-recover} supplies a positive solution of the original
Schr\"odinger system.

\subsection{Second case: no ordered iterate occurs}
Assume instead that for every $n$, the set on which $H_n'>H_n$ has positive
measure.  The monotonicity in \eqref{eq:fortet-iterate-monotone} gives
pointwise limits
\[
  H=\lim_{n\to\infty}H_n,
  \qquad
  H''=\lim_{n\to\infty}H_n''.
\]
The floors $1/n$ disappear in the limit, and the identities
$H_{n+1}=M(H_n'',1/(n+1))$ give $H=H''$.  Passing through the two
nonnegative integrals defining $Q$ yields a limiting image $H'=Q(H)$ and
the source obtains
\begin{equation}
  H\le H'. \tag{15}\label{eq:fortet-limit-order}
\end{equation}

The delicate point is exclusion of the zero fixed point.  Fortet argues that
if $H_n'(x_0)\downarrow0$ at one point $x_0$, then
$G(H_n,\cdot)^{-1}$ tends to zero in measure.  Splitting the integral for
$H_n'(x)$ into a finite central part and two tails, using continuity of $g$
on the central part and the integrability hypothesis on the tails, he obtains
$H_n'(x)\to0$ uniformly in $x$.  For sufficiently large $n$ this forces
$H_n'<H_n$ everywhere because $H_n\ge1/n$, contradicting the assumption of
the second case.  Hence the limit is strictly positive.

Now pass the mass identity to the limit:
\[
  \int_{J'}\frac{H'(x)}{H(x)}\omega_1(x)\,dx=1.
\]
By \eqref{eq:fortet-limit-order}, the quotient is at least one; therefore it
is one almost everywhere on $A_1$.  Consequently $H'=H$ almost everywhere,
and the regularity part of the fundamental lemma upgrades the resulting
kernel integrals to the pointwise/continuous form needed to recover
$(\varphi,\psi)$.

\begin{remark}[Compressed point in the source]
The source's zero-exclusion argument occupies pp.~92--94 and uses several
``uniformly in $x$'' statements with the tail integrals.  L\'eonard
\cite[Lemmas 5.10--5.12 and Theorem 5.6]{Leonard2019} isolates this as the
only step needing hypotheses beyond monotonicity and the mass identity.  The
modern citation clarifies the bookkeeping but does not replace Fortet's
contradiction mechanism.
\end{remark}

\subsection{Theorem I with its exact integrability condition}
\begin{theorem}[Fortet Theorem I, source p.~96]
Assume Hypotheses I and II and suppose
\begin{equation}
  \int_{J''}
  \frac{\omega_2(y)}{\displaystyle\int_{J'}g(z,y)\omega_1(z)\,dz}
  \,dy<\infty. \tag{T1}\label{eq:fortet-theorem1-integrability}
\end{equation}
Then the system \eqref{eq:fortet-S1}--\eqref{eq:fortet-S2} has a positive
solution.  The factor $\varphi$ is continuous and vanishes exactly where
$\omega_1$ vanishes.  The factor $\psi$ is Borel measurable and, up to a
null set, vanishes only where $\omega_2$ vanishes.
\end{theorem}

\begin{proof}[Assembly from the preceding construction]
Condition \eqref{eq:fortet-theorem1-integrability} is precisely the estimate
used in the second-case tail control and in showing that the candidate kernel
integrals are finite and continuous.  The clipped iteration yields a strictly
positive fixed point $h$ of $Q$ in either case.  Because $h\le1$ in Fortet's
normalization, the denominator in \eqref{eq:fortet-recover} is finite.  The
continuity estimate on source pp.~94--95 splits
$|h(x)-h(x_0)|$ into a finite-domain kernel-difference term and a uniformly
small tail term; continuity of $g$ handles the first and
\eqref{eq:fortet-theorem1-integrability} the second.  The recovered
$\varphi=\omega_1/h$ and $\psi=\omega_2/G(h,\cdot)$ have the stated zero sets
by the nonnegativity of all integrands and the slice nondegeneracy of $g$.
\end{proof}

Fortet remarks that the roles of the two marginals may be interchanged in
\eqref{eq:fortet-theorem1-integrability}.  He also observes that if $J''$ is
finite, boundedness of $g$ already makes the relevant tail conditions much
easier to satisfy.

\section{The class $(B)$ and the second existence theorem}
Fortet then isolates a condition on the kernel that implies the hypotheses needed in the first theorem locally and gives continuity of the recovered solution.
A function $g(x,y)$ is of class $(B)$ in $x$ on a finite interval $\mathcal F$ if convergence of integrals
\[
  \int g(x,y)f(y)\,dy
\]
which is finite a.e. on an open set containing $\mathcal F$ is in fact uniform for $x\in\mathcal F$.  The point of this definition is to upgrade almost-everywhere information to continuity in $x$.

Fortet gives sufficient conditions:
\begin{enumerate}[leftmargin=2em]
\item if $g$ is bounded below by a positive number independent of $x$ and $y$ on the finite window, then $g$ is class $(B)$;
\item more generally, if there is a uniform ratio bound
\[
  \frac{g(x_1,y)}{g(x_2,y)}<\beta
\]
for $x_1,x_2$ in the finite window and all relevant $y$, then $g$ is class $(B)$;
\item if $J'=J''=\R$ and $g(x,y)=U(x-y)$ with $U$ eventually monotone on both tails, then $g$ is class $(B)$ in both variables.
\end{enumerate}
The Gaussian kernel in Bernstein's case satisfies this tail monotonicity condition.

\subsection{Theorem II}
Let $A_1=\{x:\omega_1(x)>0\}$.  Fortet fixes a compact interval $\mathcal F\subset A_1$ and chooses an auxiliary positive continuous density $p(x)$ on $\mathcal F$ with
\[
  \int_{\mathcal F}p(x)\,dx=1,
\]
plus integrability inequalities making Theorem I applicable to the restricted system.  Solving the restricted system gives a local positive pair.  He then sets up a second fixed-point equation on $A_1$ and repeats the monotone argument.  The class $(B)$ property makes the kernel integrals uniformly convergent on compact subintervals of $A_1$, and therefore the recovered $h$ and hence $\varphi$ are continuous on $A_1$.

\begin{theorem}[Fortet Theorem II, source pp.~101--102]
Under Hypotheses I and II, if $g$ is of class $(B)$ in the $x$ variable, then
the system has a solution with
\begin{enumerate}[leftmargin=2em]
\item $\varphi(x)=0$ when $\omega_1(x)=0$;
\item $\varphi(x)>0$ and continuous when $\omega_1(x)>0$;
\item $\psi(y)\ge0$ is Borel measurable and vanishes, modulo null sets, only where $\omega_2(y)=0$.
\end{enumerate}
\end{theorem}

Fortet then notes that if $\omega_1>0$ on $A_1$ and $\omega_2>0$ on $A_2$, solving the restricted system on $A_1\times A_2$ and extending by zero outside those sets gives a solution on the original domains.

\section{Uniqueness}
Fortet proves uniqueness among positive Borel solutions, modulo the scalar gauge.
Let $(\varphi_1,\psi_1)$ and $(\varphi_2,\psi_2)$ be two positive solutions.  On $A_1=\{\omega_1>0\}$ set
\[
  h_i(x)=\frac{\omega_1(x)}{\varphi_i(x)}.
\]
Each $h_i$ satisfies the same fixed-point equation
\[
  h_i(x)=\int_{J''}g(x,y)
  \frac{\omega_2(y)}{G(h_i,y)}\,dy.
\]
Let
\[
  h(x)=M(h_1(x),h_2(x))=\max\{h_1(x),h_2(x)\}.
\]
Then $G(h,y)$ is finite a.e. on $A_2=\{\omega_2>0\}$, because $h\ge h_i$ implies $1/h\le1/h_i$ and the original denominators are finite.  Define
\[
  h'(x)=Q(h)(x)=\int g(x,y)\frac{\omega_2(y)}{G(h,y)}\,dy.
\]
The mass identity gives
\[
  \int\frac{h'}{h}\omega_1\,dx=1. \tag{U1}\label{eq:fortet-uniq-mass}
\]
Since $h\ge h_i$, we have $G(h,y)\le G(h_i,y)$ and therefore
\[
  Q(h)(x)=h'(x)\ge Q(h_i)(x)=h_i(x)
\]
for $i=1,2$.  Thus
\[
  h'\ge h.
\]
Together with \eqref{eq:fortet-uniq-mass}, the comparison lemma forces
\[
  h'=h \quad\hbox{a.e. on }A_1.
\]
From $h=h_1$ on the subset where $h_1\ge h_2$ and the equality of denominators, Fortet derives
\[
  G(h,y)=G(h_1,y)\quad\hbox{a.e. on }A_2,
\]
and hence the integrands force $h=h_1$ a.e.  Similarly $h=h_2$ a.e.  Therefore
\[
  h_1=h_2\quad\hbox{a.e. on }A_1.
\]
Undoing the substitution $h_i=\omega_1/\varphi_i$ gives $\varphi_1=\varphi_2$ after the chosen gauge normalization.  Without a fixed gauge, the result is uniqueness up to multiplication of one factor by a positive constant and the other by its reciprocal.

\begin{theorem}[Fortet Theorem III, source p.~104]
Under Hypotheses I and II, the Schr\"odinger system has at most one positive
Borel solution, where solutions differing by the scalar gauge
$(\varphi,\psi)\mapsto(c\varphi,c^{-1}\psi)$ are identified.
\end{theorem}


\section{Fortet's three closing remarks}
Fortet appends three remarks after Theorem III.  They are not used in the
existence proof, but they clarify the exact scope of the result and the role
of positivity.

\subsection*{Remark I: continuity of the marginals can be weakened}
The arguments of Sections 2--4 continue, with only minor changes in the
statements, if $\omega_1$ and $\omega_2$ are assumed merely Borel measurable
rather than continuous.  What is lost is the automatic continuity conclusion
for the recovered factor; the order, truncation, and mass-identity arguments
remain measure-theoretic.

\subsection*{Remark II: bounded solutions on a finite second interval}
Assume $\omega_1$ is continuous, $J''$ is finite, and restrict attention to
solutions for which both factors are bounded.  Then
\[
  x\longmapsto \int_{J''}g(x,y)\psi(y)\,dy
\]
is continuous, and therefore so is
\[
  \varphi(x)=
  \frac{\omega_1(x)}{\int_{J''}g(x,y)\psi(y)\,dy}.
\]
If $\omega_1$ never vanishes, a continuous $\varphi$ could change sign only by
passing through zero.  At such a point the displayed quotient would require
the denominator to become infinite, contradicting boundedness of $\psi$ and
finiteness of $J''$.  Thus, under these extra hypotheses, every bounded
solution has a fixed sign and can be put in the positive class by the scalar
gauge.  Fortet concludes that Theorem I then supplies the only bounded
solution, not merely the only positive one.

\subsection*{Remark III: zeros of the marginals permit signed solutions}
Fortet gives an explicit warning that uniqueness among positive solutions does
not imply uniqueness among all signed solutions when a marginal vanishes.
Take
\[
  J'=J''=[0,3\pi]
\]
and define
\[
  \omega_1(x)=
  \begin{cases}
    \frac14\sin x,&0\le x\le\pi,\\
    0,&\pi\le x\le2\pi,\\
    \frac14\sin x,&2\pi\le x\le3\pi,
  \end{cases}
  \qquad
  \varphi(x)=
  \begin{cases}
    \sin x,&0\le x\le\pi,\\
    0,&\pi\le x\le2\pi,\\
    -\sin x,&2\pi\le x\le3\pi.
  \end{cases}
\]
Choose a continuous kernel $g$ whose values on the two active $x$-blocks are
\[
  g(x,y)=
  \begin{cases}
    \frac14,&0\le y\le\pi,\\
    \frac18,&2\pi\le y\le3\pi,
  \end{cases}
  \quad 0\le x\le\pi,
\]
and
\[
  g(x,y)=
  \begin{cases}
    \frac18,&0\le y\le\pi,\\
    \frac14,&2\pi\le y\le3\pi,
  \end{cases}
  \quad 2\pi\le x\le3\pi,
\]
with an arbitrary continuous interpolation elsewhere.  Direct substitution
shows that $\varphi$ solves the reduced integral equation while changing sign.
The example is deliberately elementary: the zero strip in $\omega_1$ allows
the two active components to carry opposite signs without violating the
marginal equation.

\section{Bernstein's Gaussian case}
Fortet's motivating special case, attributed to Bernstein, has
\[
  J'=J''=\R,
  \qquad
  g(x,y)=\frac{1}{a\sqrt\pi}\exp\left(-\frac{(x-y)^2}{a^2}\right),
  \qquad a>0,
\]
with continuous marginals.  The Gaussian kernel is a function of $x-y$ and is monotone on each tail, so it is class $(B)$ in both variables.  Therefore Theorems II and III imply:

\begin{distilled}[Fortet's Bernstein conclusion]
In the Gaussian Bernstein case, the Schrodinger system admits a positive continuous solution, and this solution is the unique positive Borel solution modulo the scalar gauge.
\end{distilled}

\section{Structural commentary}
\begin{commentarynotes}
\item The fixed-point operator is the continuum analogue of composing the two quotient updates in matrix scaling.
\item The mass identity \eqref{eq:fortet-mass-identity} is the device that removes residual scalar freedom from one-sided comparison arguments.
\item Fortet's proof is an order-and-compactness argument, not a Hilbert-metric contraction proof; its distinctive ingredients are truncation, monotone iteration, the mass identity, and local uniform compactness.
\item The finite-dimensional specialization retains the same algebraic skeleton: denominator map $\to$ update map $\to$ mass identity, and max--min comparison.
\end{commentarynotes}

\section{Source-fidelity and correction notes}
\begin{commentarynotes}
\item The exact three-sequence clipping scheme on source p.~90 is retained.
Replacing it by untruncated iteration $H_{n+1}=Q(H_n)$ would not be Fortet's
proof and would reintroduce the zero and infinity fixed points that the floors
and ceilings are designed to avoid.
\item Fortet invokes a Beppo--Levi theorem ``completed by de la Vall\'ee
Poussin'' when passing the two truncation limits in the mass identity.  In
modern language, Tonelli and monotone convergence justify the nonnegative
integrals; any additional uniform-integrability passage is stated separately.
\item The continuity and strict-positivity proof in the second case is terse.
The reconstruction preserves its finite-part/tail-splitting argument and cites
L\'eonard only for a modern organization of the same step.
\item Theorem III is uniqueness only after quotienting by the scalar gauge.
Fortet fixes the gauge by choosing a point $x_0$ at which the two first factors
are equal before applying the maximum argument.
\item The title and formulas in the 1940 paper use ``Schr\"odinger'' with
historical typography and no modern measure-kernel notation.  The normalized
notation here changes presentation, not the order-theoretic proof.
\end{commentarynotes}

\section*{Checklist of details retained}
\begin{itemize}[leftmargin=2em]
\item The two-equation system and gauge equivalence.
\item The reduction from $(\varphi,\psi)$ to a single equation $h=Q(h)$.
\item Class $(C)$ and why it is used.
\item The truncated-denominator construction and monotone convergence.
\item The normalization identity $\int Q(H)/H\,d\omega_1=1$.
\item The comparison lemma turning $Q(H)\le H$ or $Q(H)\ge H$ into equality.
\item The two-case successive-approximation proof.
\item The class $(B)$ local-uniform-convergence condition and the Gaussian/Bernstein application.
\item The uniqueness proof using $h=\max(h_1,h_2)$.
\item Fortet's three closing remarks, including the explicit signed-solution example.
\end{itemize}

\input{references_manual}
\end{document}
