\input{preamble}
\title{Absolutely Continuous Changes of Measure for Wiener and Diffusion Processes}
\author{}
\date{Girsanov (1960): Fundamental Theorems 1--2 and Lemmas 1--7}
\begin{document}
\maketitle

\section*{Source map and scope}
\begin{description}[style=nextline,leftmargin=3.2cm,labelwidth=3cm]
\item[Primary source] \texttt{\detokenize{girsanov1960.pdf}}
\item[Coverage] Sections 1--3: the conditional-moment definition of Wiener processes, the exponential density, Lemmas 1--6 and Fundamental Theorem 1, then Lemma 7 and Fundamental Theorem 2 for diffusion-type processes.
\item[Source anchors] equations (1.1)--(1.4), (2.1)--(2.11), (3.1), Fundamental Theorems 1--2, Lemmas 1--7, and the remarks following Lemma 6 and Theorem 2.
\end{description}

This reconstruction preserves the source's proof architecture: first prove the theorem for bounded
integrands, then approximate by cutoffs, use Fatou and convergence of stochastic integrals, verify the
conditional first and second moments of the shifted noise, and only afterward transfer the result to
path-dependent diffusion equations.  Modern stochastic-exponential notation is used, but the proof is
not replaced by a later Novikov- or Kazamaki-based theorem.

\section{The historical setup}
Girsanov writes an \(n\)-dimensional process as \(X=\{x(t,\omega),P\}\), \(0\le t\le1\), on a
filtered measurable space \((\Omega,\mathcal F,(\mathcal F_t),P)\).  A Wiener process
\(\xi=(\xi_1,\dots,\xi_n)\) is characterized by continuity and the conditional identities, for
\(0\le s<t\le1\),
\begin{align}
 E_P[\xi_i(t)-\xi_i(s)\mid\mathcal F_s]&=0,\label{eq:girs-cond1}\\
 E_P[(\xi_i(t)-\xi_i(s))(\xi_j(t)-\xi_j(s))\mid\mathcal F_s]
 &=\delta_{ij}(t-s).\label{eq:girs-cond2}
\end{align}
The source invokes L\'evy's characterization to identify a continuous process satisfying these
relations as Wiener.

An It\^o process relative to \(\xi\) is written
\begin{equation}\label{eq:girs-ito}
 x(t)=x(0)+\int_0^t a(s)\,ds+\int_0^t q(s)\,d\xi(s),
\end{equation}
where \(a\) is vector-valued and \(q\) is matrix-valued.  The historical assumptions
\((A_0)\)--\((A_3)\) impose adapted measurability and pathwise square-integrability; when an
inverse matrix is used, nondegeneracy and corresponding integrability are also assumed.  The paper
works directly with these hypotheses rather than collecting them into a single modern martingale
criterion.

\section{The exponential substitution}
Let \(\varphi=(\varphi_1,\dots,\varphi_n)\) be adapted and square-integrable along almost every
path.  Set
\begin{align}
 M_t&=\int_0^t\varphi(s)\cdot d\xi(s),\label{eq:girs-M}\\
 A_t&=\int_0^t|\varphi(s)|^2\,ds,\label{eq:girs-A}\\
 Z_t&=\exp\{M_t-\tfrac12A_t\}.\label{eq:girs-Z}
\end{align}
The source assumes the terminal normalization
\begin{equation}\label{eq:girs-normalization}
 E_P Z_1=1.
\end{equation}
It then defines
\[
 \widetilde P(d\omega)=Z_1(\omega)P(d\omega)
\]
and the shifted noise
\begin{equation}\label{eq:girs-shifted}
 \widetilde\xi(t)=\xi(t)-\int_0^t\varphi(s)\,ds.
\end{equation}
The sign in \eqref{eq:girs-shifted} is tied to the positive sign in the stochastic integral in
\eqref{eq:girs-Z}.

\section{It\^o's formula and the bounded estimates}
Applying It\^o's formula to \(e^u\) and \(u=M_t-\tfrac12A_t\) gives
\begin{equation}\label{eq:girs-dZ}
 Z_t=1+\int_0^t Z_s\varphi(s)\cdot d\xi(s).
\end{equation}
For bounded \(\varphi\), this is a square-integrable martingale.  The source does not simply cite
that conclusion: Lemma 1 obtains explicit exponential-moment bounds by stopping the exponential,
applying It\^o's formula to powers, taking expectations, and using a Gronwall estimate.  In modern
notation, if \(|\varphi|\le c\), then for \(r>1\)
\[
 E_P[Z_t^r]
 \le \exp\!\left(\frac{r(r-1)}2c^2t\right).
\]
The precise constants in the translated scan are expressed through the source's matrix/vector norms;
the role of the estimate is uniform integrability of bounded approximations.

\section{Cutoffs and convergence: Lemmas 2--5}
Let \(W_N\) be the source's cutoff, equal to one on a bounded interval and zero outside a slightly
larger interval.  The truncated integrands are
\[
 \varphi^{(N)}(t)=W_N\!\left(\int_0^t|\varphi(s)|^2ds\right)\varphi(t),
\]
or equivalent coordinatewise/pathwise cutoffs used in individual lemmas.  Write
\(M^{(N)},A^{(N)},Z^{(N)}\) for the corresponding quantities.

\paragraph{Lemma 2: convergence of stochastic integrals.}
If \(q^{(N)}\to q\) in the source's pathwise \(L^2\) sense and the square integrals are uniformly
controlled, then
\[
 \int_0^tq^{(N)}(s)d\xi(s)
 \longrightarrow \int_0^tq(s)d\xi(s)
\]
in probability, and along a subsequence almost surely.  The proof uses the It\^o isometry for bounded
approximants and then truncates the exceptional set where the quadratic integral is large.

\paragraph{Lemma 3: weighted integral identity.}
For bounded adapted \(q\) and bounded \(\mathcal F_s\)-measurable \(F\), the product formula applied
to \(Z_t\) and the stochastic integral from \(s\) to \(t\) yields the cancellation identity
\begin{equation}\label{eq:girs-weighted-first}
 E_P\!\left[FZ_t\left(\xi_i(t)-\xi_i(s)-
      \int_s^t\varphi_i(u)du\right)\right]=0.
\end{equation}
Applying the product rule once more to pairs of shifted increments gives
\begin{multline}\label{eq:girs-weighted-second}
 E_P\!\left[FZ_t
  \left(\widetilde\xi_i(t)-\widetilde\xi_i(s)\right)
  \left(\widetilde\xi_j(t)-\widetilde\xi_j(s)\right)\right]\\
 =\delta_{ij}(t-s)E_P[FZ_s].
\end{multline}
These are the source's direct replacements for an abstract semimartingale-characteristics theorem.

\paragraph{Lemma 4: intermediate-time density.}
The function denoted \(g(t,\omega)\) in the source is the terminal exponential conditional on
\(\mathcal F_t\):
\[
 g_t=E_P[Z_1\mid\mathcal F_t].
\]
For bounded integrands, \eqref{eq:girs-dZ} and the normalization show \(g_t=Z_t\).  The lemma
establishes the integrability and conditional-density relation needed to replace terminal weights
\(Z_1\) by current weights \(Z_t\) in \eqref{eq:girs-weighted-first}--\eqref{eq:girs-weighted-second}.

\paragraph{Lemma 5: passage to the normalized unbounded case.}
Apply the bounded results to \(\varphi^{(N)}\).  The cutoffs eventually agree with \(\varphi\) on
each path because \(A_1<\infty\) almost surely.  Lemmas 1--2 give convergence of the exponentials
and stochastic integrals.  Fatou gives
\(E_P Z_1\le\liminf E_P Z^{(N)}_1=1\); the assumed equality
\eqref{eq:girs-normalization} rules out loss of mass.  Scheff\'e's argument for nonnegative densities
with equal integrals then yields \(L^1\)-convergence along the required approximation, allowing the
weighted moment identities to pass to the limit.

\section{Fundamental Theorem 1}
\begin{theorem}[Girsanov's Fundamental Theorem 1]
Let \(\xi\) be Wiener under \(P\), let \(x\) be the It\^o process
\eqref{eq:girs-ito}, and let \(\varphi\) satisfy the source assumptions and
\(E_PZ_1=1\).  Under \(\widetilde P=Z_1P\), the process
\[
 \widetilde\xi(t)=\xi(t)-\int_0^t\varphi(s)ds
\]
is Wiener.  Relative to \(\widetilde\xi\), the same process \(x\) has diffusion matrix \(q\) and
drift
\begin{equation}\label{eq:girs-drift-shift}
 \widetilde a(t)=a(t)+q(t)\varphi(t).
\end{equation}
\end{theorem}

\begin{proof}[Source proof]
For bounded \(\varphi\), \eqref{eq:girs-weighted-first} and
\eqref{eq:girs-weighted-second}, together with
\(E_{\widetilde P}[Y]=E_P[Z_1Y]\), prove the conditional moment relations
\eqref{eq:girs-cond1}--\eqref{eq:girs-cond2} for \(\widetilde\xi\).  Continuity is unchanged, so
L\'evy's criterion makes \(\widetilde\xi\) Wiener.  Lemmas 2--5 extend the identities to the general
normalized exponential.

For the drift statement, write
\[
 d\xi(t)=d\widetilde\xi(t)+\varphi(t)dt
\]
and substitute into \eqref{eq:girs-ito}.  The only nonalgebraic point is that a stochastic integral
with respect to the old Wiener process has the asserted decomposition under the new measure.  This
is supplied by Lemma 6 below.
\end{proof}

\section{Lemma 6: transformation of stochastic integrals}
Let \(S=\{\xi,P\}\) and \(\widetilde S=\{\widetilde\xi,\widetilde P\}\) be the two Wiener systems of
Theorem 1, and let \(Q(t,\omega)\) satisfy the source's adaptedness and square-integrability
conditions.  Lemma 6 proves
\begin{equation}\label{eq:girs-integral-transform}
 \int_0^tQ(s)d\xi(s)
 =\int_0^tQ(s)d\widetilde\xi(s)+
   \int_0^tQ(s)\varphi(s)ds,
 \qquad \widetilde P\text{-a.s.}
\end{equation}
The proof begins with bounded elementary integrands, where it is the defining Riemann-sum identity.
It then approximates \(Q\) by bounded processes and uses Lemma 2 under both measures.  This is why
the paper treats Lemma 6 separately instead of regarding \eqref{eq:girs-integral-transform} as a
formal substitution.

The remark following Lemma 6 observes that a nontrivial issue is the terminal normalization
\(E_PZ_1=1\).  Local exponential martingales need not preserve mass.  The source gives sufficient
boundedness-type hypotheses; modern Novikov and Kazamaki criteria are later refinements, not part of
the 1960 proof.

\section{Diffusion-type processes}
Section 3 considers path-dependent coefficients.  Let \(C_n=C([0,1];\mathbb R^n)\), and write
\(x^t\) for the path stopped at \(t\).  A diffusion-type process satisfies
\begin{equation}\label{eq:girs-path-sde}
 dx(t)=A(t,x^t)dt+B(t,x^t)d\xi(t).
\end{equation}
The coefficients are nonanticipating: their value at time \(t\) depends only on \(x^t\).

\paragraph{Lemma 7: recovery of the driving noise.}
Under nondegeneracy, solve for a matrix \(P(t,x^t)\) satisfying
\[
 P(t,x^t)B(t,x^t)=I.
\]
Then the process
\begin{equation}\label{eq:girs-recovered-noise}
 \xi(t)=\int_0^tP(s,x^s)dx(s)-
        \int_0^tP(s,x^s)A(s,x^s)ds
\end{equation}
is Wiener.  The proof applies It\^o's formula and the conditional-characteristic identities.  The
lemma makes the density functional measurable in the observed path \(x\), rather than in an
auxiliary noise coordinate.

Given a path functional \(\varphi(t,x^t)\), define
\begin{equation}\label{eq:girs-Zpath}
 \mathcal Z(x)=
 \int_0^1\varphi(t,x^t)\cdot d\xi(t)
 -\frac12\int_0^1|\varphi(t,x^t)|^2dt,
\end{equation}
where \(\xi\) is reconstructed by \eqref{eq:girs-recovered-noise}.  In the source this is rewritten,
using \eqref{eq:girs-path-sde}, as a stochastic line integral in \(x\) plus finite-variation terms.
Consequently the Radon--Nikodym derivative can be expressed as a functional on path space:
\[
 \frac{d\widetilde\mu}{d\mu}(x)=e^{\mathcal Z(x)}.
\]

\section{Fundamental Theorem 2}
\begin{theorem}[Girsanov's Fundamental Theorem 2]
Suppose \(X\) is a diffusion-type process with diffusion matrix \(B(t,x^t)\) and drift
\[
 A(t,x^t)+B(t,x^t)\varphi(t,x^t).
\]
Assume the reference equation \eqref{eq:girs-path-sde} has a unique solution, the coefficient
hypotheses of the source hold, and \(|\varphi(t,x^t)|\) is bounded by a nondecreasing function of
\(\sup_{s\le t}|x(s)|\).  If
\[
 \widetilde P(d\omega)=e^{\mathcal Z(x(\cdot,\omega))}P(d\omega),
\]
then under \(\widetilde P\) the shifted process
\[
 \widetilde\xi(t)=\xi(t)-\int_0^t\varphi(s,x^s)ds
\]
is Wiener and \(X\) has drift \(A(t,x^t)\) and diffusion matrix \(B(t,x^t)\).  The induced path law
\(\widetilde\mu\) is absolutely continuous with respect to \(\mu\), with density
\(e^{\mathcal Z(x)}\).
\end{theorem}

\begin{proof}[Source proof]
Stop the process when \(\sup_{s\le t}|x(s)|\) reaches \(N\).  On the stopped set the growth bound
makes \(\varphi\) bounded, so Theorem 1 applies.  Lemma 7 expresses the exponential as a path
functional.  Letting \(N\to\infty\) gives the unrestricted result.  Uniqueness of the reference SDE
identifies the transformed law: after the drift change, the coordinate process satisfies the same
equation as the reference diffusion, so the laws coincide.
\end{proof}

The final remarks emphasize that uniqueness of the diffusion equation is indispensable to this law
identification.  The paper mentions uniform Lipschitz hypotheses as a sufficient condition, and
weaker H\"older/nondegeneracy conditions for time-homogeneous Markov diffusions.

\section{Later entropy consequence}
The 1960 article proves absolute continuity and drift transformation, not an entropy theorem.
A standard later consequence, included here with explicit provenance, is the control-energy identity
\cite{Leonard2012Girsanov,RogersWilliams2000}.  If \(Q=Z_TP\) and under \(Q\)
\(\widetilde W_t=W_t-\int_0^tu_sds\) is Brownian, then
\[
 \log\frac{dQ}{dP}
 =\int_0^Tu_t\cdot d\widetilde W_t+
   \frac12\int_0^T|u_t|^2dt.
\]
Under the integrability assumptions that make the first term a true \(Q\)-martingale,
\begin{equation}\label{eq:girs-entropy-energy}
 H(Q\mid P)=\frac12E_Q\int_0^T|u_t|^2dt.
\end{equation}
This is not presented as part of Girsanov's original theorem; it is the modern corollary used in
Schr\"odinger-bridge and stochastic-control arguments.

\section{Source issues and normalization}
\begin{enumerate}[leftmargin=2em]
\item The English scan is typographically noisy: several symbols in the statement of Theorem 1
are partially lost.  Equations \eqref{eq:girs-Z}--\eqref{eq:girs-drift-shift} are fixed by the
surrounding proof and the source's substitution formula.
\item The source assumes \(E_PZ_1=1\); it does not claim that pathwise square-integrability alone
implies this normalization.
\item Theorem 2 uses uniqueness of the path-dependent SDE to identify laws.  Theorem 1 alone only
identifies the transformed local characteristics.
\item Modern terms such as ``stochastic exponential,'' ``uniform integrability,'' and ``Scheff\'e
argument'' describe the exact operations carried out in the source, but were not its terminology.
\end{enumerate}

\input{references_manual}
\end{document}
