\input{preamble}
\title{The Schr\"odinger Problem, $(f,g)$-Transforms, and the Slow-Noise Limit}
\author{}
\date{L\'eonard (2014): Sections 2--5}
\begin{document}
\maketitle

\section*{Source map and scope}
\begin{description}[style=nextline,leftmargin=3.2cm,labelwidth=3cm]
\item[Primary source] \texttt{\detokenize{leonard-2013_survey-schrodinger-problem-optimal-transport__arXiv-1308.0215.pdf}}
\item[Coverage] Definitions 2.1--2.2, Proposition 2.3, Lemma 2.4, Propositions 2.5--2.7, Theorems 2.8--2.9, Proposition 2.10 and Claim 2.11, Theorem 2.12, Definition 3.2, Theorems 3.3--3.4, Propositions 4.1--4.2, the Section 5 Laplace-principle derivation, and Theorem 5.2.
\item[Source anchors] equations (2.2)--(2.12), (3.2)--(3.9), (4.1)--(4.16), and (5.1)--(5.3).
\end{description}

This note follows the survey's order.  Proofs supplied in the survey are expanded.  Results that the
survey imports from earlier work---notably the existence of additive Schr\"odinger potentials in full
generality and the rigorous Gamma-convergence theorems---retain that citation-dependent status.

\section{Sigma-finite entropy and the two Schr\"odinger problems}
Let \(\Omega\) be a Polish path space over a Polish state space \(\mathcal X\), with canonical
coordinate \(X_t\).  The reference \(R\in\mathcal M_+(\Omega)\) may be sigma-finite.  L\'eonard
defines relative entropy with respect to such an \(R\) by choosing a nonnegative weight \(W\) for
which
\[
 z_W=\int e^{-W}dR<\infty,
 \qquad
 R_W=z_W^{-1}e^{-W}R\in\mathcal P(\Omega),
\]
and setting
\begin{equation}\label{eq:leo-sigmafinite-entropy}
 H(P\mid R)=H(P\mid R_W)-\int WdP-\log z_W.
\end{equation}
The value is independent of the admissible weight.

For prescribed \(\mu_0,\mu_1\in\mathcal P(\mathcal X)\), define
\begin{align}
 (S_{\rm dyn})\quad&	ext{minimize }H(P\mid R)
   \text{ over }P\in\mathcal P(\Omega),\ P_0=\mu_0,\ P_1=\mu_1,
   \label{eq:leo-Sdyn}\\
 (S)\quad&	ext{minimize }H(\pi\mid R_{01})
   \text{ over }\pi\in\mathcal P(\mathcal X^2),\ \pi_0=\mu_0,\ \pi_1=\mu_1.
   \label{eq:leo-S}
\end{align}

\section{Proposition 2.3: dynamic--static equivalence}
Disintegrate
\[
 R(d\omega)=R^{xy}(d\omega)R_{01}(dx,dy),
 \qquad
 P(d\omega)=P^{xy}(d\omega)P_{01}(dx,dy).
\]
The entropy chain rule on Polish spaces gives
\begin{equation}\label{eq:leo-chain}
 H(P\mid R)=H(P_{01}\mid R_{01})+
 \int H(P^{xy}\mid R^{xy})P_{01}(dx,dy).
\end{equation}
Thus \(H(P\mid R)\ge H(P_{01}\mid R_{01})\), with equality at finite entropy exactly when
\(P^{xy}=R^{xy}\) for \(P_{01}\)-almost every endpoint pair.

\begin{proposition}[Proposition 2.3]
The two problems have at most one minimizer.  Their infima agree.  If \(\widehat\pi\) solves the
static problem, then
\begin{equation}\label{eq:leo-bridge-mixture}
 \widehat P(\cdot)=\int R^{xy}(\cdot)\widehat\pi(dx,dy)
\end{equation}
solves the dynamic problem.  Conversely, the endpoint law of a dynamic minimizer is the static
minimizer and the dynamic minimizer shares all endpoint bridges with \(R\).
\end{proposition}

\section{Lemma 2.4 and Proposition 2.5: existence}
Let \(B:\mathcal X\to[0,\infty)\) satisfy
\[
 \int e^{-B(x)-B(y)}R_{01}(dx,dy)<\infty,
 \qquad
 \int Bd\mu_0+\int Bd\mu_1<\infty.
\]
Then the constraint set \(\Pi(\mu_0,\mu_1)\) is weakly compact by tightness of its fixed marginals.
After tilting \(R_{01}\) to a probability, the entropy differs by a constant on this constraint set;
therefore it is lower semicontinuous and bounded below.  Lemma 2.4 concludes:
\begin{equation}\label{eq:leo-existence-criterion}
 \text{a minimizer exists}
 \quad\Longleftrightarrow\quad
 \exists\pi^o\in\Pi(\mu_0,\mu_1):H(\pi^o\mid R_{01})<\infty.
\end{equation}

Suppose \(R_0=R_1=m\).  Proposition 2.5 first notes the necessary conditions
\(H(\mu_0\mid m),H(\mu_1\mid m)<\infty\), by data processing.  For sufficiency, assume
\[
 R_{01}(dx,dy)\ge e^{-A(x)-A(y)}m(dx)m(dy)
\]
and suitable integrability of \(A+B\) under both marginals.  Testing
\eqref{eq:leo-existence-criterion} with \(\mu_0\otimes\mu_1\) gives finite entropy relative to
\(R_{01}\).  Exponential integrability of \(A+B\) under \(m\), together with finite marginal
entropy and the entropy variational formula, implies the required first moments.

\section{Duality and additive potentials}
For a weight \(B\), let \(C_B(\mathcal X)\) be the continuous functions growing at most like \(B\).
The static dual is
\begin{equation}\label{eq:leo-dual}
 \sup_{\varphi,\psi\in C_B}
 \left\{
  \int\varphi d\mu_0+\int\psi d\mu_1
  -\log\int e^{\varphi(x)+\psi(y)}R_{01}(dx,dy)
 \right\}.
\end{equation}
The survey states the dual equality and explains that if the dual is attained, equality in the
entropy variational formula forces
\[
 \frac{d\widehat\pi}{dR_{01}}=e^{\varphi\oplus\psi}.
\]
In general, existence of separately measurable potentials is subtler.  Proposition 2.6 uses an
``internal constraint'' condition to ensure positivity and additive form.  Proposition 2.7 uses
\(R_{01}\ll R_0\otimes R_1\) to obtain separately measurable versions.  The alternative domination
\(R_0\otimes R_1\ll R_{01}\) also makes an \(R_{01}\)-measurable sum product-measurable.

\section{Theorem 2.8: the Schr\"odinger system}
Assume the seven hypotheses listed by the survey: common endpoint reference \(m\), a lower kernel
bound with weight \(A\), an integrable tilt with weight \(B\), one of the two product-domination
conditions, finite marginal entropies, integrability of \(A+B\), and internality of the constraint.
Then the unique static minimizer is
\begin{equation}\label{eq:leo-static-factor}
 \widehat\pi(dx,dy)=f_0(x)g_1(y)R_{01}(dx,dy),
\end{equation}
where \(f_0,g_1>0\) are \(m\)-measurable and solve
\begin{align}
 f_0(x)E_R[g_1(X_1)\mid X_0=x]&=\frac{d\mu_0}{dm}(x),\label{eq:leo-system1}\\
 g_1(y)E_R[f_0(X_0)\mid X_1=y]&=\frac{d\mu_1}{dm}(y).\label{eq:leo-system2}
\end{align}
The factors need not be integrable individually; nonnegative conditional integration suffices.  They
are determined only up to \((f_0,g_1)\mapsto(cf_0,c^{-1}g_1)\).

Theorem 2.9 combines \eqref{eq:leo-bridge-mixture} and \eqref{eq:leo-static-factor}:
\begin{equation}\label{eq:leo-dynamic-factor}
 \widehat P=f_0(X_0)g_1(X_1)R.
\end{equation}
A direct entropy proof of optimality is useful.  For every competitor with the same endpoints,
\[
 \int\log(f_0(X_0)g_1(X_1))dP
 =\int\log(f_0(X_0)g_1(X_1))d\widehat P.
\]
Therefore
\[
 H(P\mid R)=H(P\mid\widehat P)+H(\widehat P\mid R),
\]
which proves uniqueness.

\section{Proposition 2.10 and Claim 2.11: Markov inheritance}
Assume \(R\) is Markov.  Fix \(t\), a time marginal \(\eta\), and prescribed conditional laws of
the past and future given \(X_t=z\).  Claim 2.11 says that among all path laws with those data, the
unique entropy minimizer relative to \(R\) is the conditionally independent gluing
\[
 P^*(d\omega)=\int Q^{z}_{<}(d\omega_{[0,t]})
                  Q^{z}_{>}(d\omega_{[t,1]})\eta(dz).
\]
The proof disintegrates entropy at \(X_t\), then disintegrates each conditional path law into past and
future.  Since \(R\) is conditionally independent, convexity of entropy gives
\[
 H\!\left(\int K_u\alpha(du)\middle|R_{>}^{z}\right)
 \le\int H(K_u\mid R_{>}^{z})\alpha(du),
\]
with equality exactly for the independent gluing.  If \(\widehat P\) were not Markov, replacing its
conditional past/future coupling by this gluing would preserve every one-time marginal, hence the
endpoint constraints, and strictly lower entropy.  Thus Proposition 2.10: the Schr\"odinger bridge of
a Markov reference is Markov.

Theorem 2.12 is the survey's more efficient Markov theorem.  Under its kernel and entropy
hypotheses, the minimizer has the factor form \eqref{eq:leo-dynamic-factor} without the same
internality assumption required in Theorem 2.8; positivity and the Markov structure allow the
potentials to be constructed on the effective support.

\section{$(f,g)$-transforms: Definition 3.2 and Theorems 3.3--3.4}
For nonnegative measurable \(f_0,g_1\) with
\(E_R[f_0(X_0)g_1(X_1)]<\infty\), define
\[
 P=f_0(X_0)g_1(X_1)R.
\]
Set
\begin{equation}\label{eq:leo-ftgt}
 f_t(z)=E_R[f_0(X_0)\mid X_t=z],
 \qquad
 g_t(z)=E_R[g_1(X_1)\mid X_t=z].
\end{equation}
Theorem 3.3 proves that the transform is Markov and gives the tilted forward and backward kernels:
\begin{align}
 P(X_t\in dy\mid X_s=x)
 &=\frac{g_t(y)}{g_s(x)}R(X_t\in dy\mid X_s=x),\label{eq:leo-forward}\\
 P(X_s\in dx\mid X_t=y)
 &=\frac{f_s(x)}{f_t(y)}R(X_s\in dx\mid X_t=y).\label{eq:leo-backward}
\end{align}
The marginal density is
\begin{equation}\label{eq:leo-marginal}
 \frac{dP_t}{dR_t}=f_tg_t.
\end{equation}
These formulas follow by inserting the endpoint factor and using the conditional independence of
past and future under \(R\).

Theorem 3.4, called the Euclidean analogue of Born's formula, states that the time marginal of the
Schr\"odinger bridge is the product of the forward and backward propagated factors:
\[
 \mu_t(dz)=f_t(z)g_t(z)m(dz)
\]
when the reference time marginal is the reversible measure \(m\).  Writing
\(\varphi_t=\log f_t\), \(\psi_t=\log g_t\) gives the additive potential representation
\(d\mu_t/dm=e^{\varphi_t+\psi_t}\).

\section{Continuous Brownian example: Proposition 4.1}
For reversible Brownian reference on \(\mathbb R^n\), the forward drift of the
\((f,g)\)-transform is \(\nabla\psi_t\), where \(\psi_t=\log g_t\), and
\begin{equation}\label{eq:leo-hjb}
 \partial_t\psi+\frac12\Delta\psi+\frac12|\nabla\psi|^2=0.
\end{equation}
Girsanov's formula gives, for any path law with initial law \(\mu_0\) and forward drift \(\beta\),
\begin{equation}\label{eq:leo-energy}
 H(P\mid R)=H(\mu_0\mid m)+
 E_P\int_0^1\frac12|\beta_t|^2dt.
\end{equation}
Therefore the bridge solves the minimum-action problem
\[
 \inf\int_0^1\int\frac12|v_t(x)|^2\nu_t(dx)dt
\]
subject to
\[
 (\partial_t-\tfrac12\Delta)\nu+
 \nabla\cdot(\nu v)=0,
 \qquad \nu_0=\mu_0,\ \nu_1=\mu_1.
\]
The survey's proof completes the square using \eqref{eq:leo-hjb}.  For a competitor \((\nu,v)\),
integrate the HJB equation against \(\nu\), use the Fokker--Planck equation, and obtain
\[
 \int\tfrac12|v|^2d\nu dt
 =\int\tfrac12|v-\nabla\psi|^2d\nu dt+
   \int\psi_1d\mu_1-\int\psi_0d\mu_0.
\]
The boundary term is fixed; equality occurs only for \(v=\nabla\psi\).

\section{Random-walk example: Proposition 4.2}
For a reversible continuous-time random walk with jump kernel \(J_x(dy)\), the transformed jump
rate is
\[
 j^\psi(t,x;y)=\exp(\psi_t(y)-\psi_t(x)).
\]
The nonlinear equation is
\[
 (\partial_t+L)\psi_t(x)+
 \int\theta(\psi_t(y)-\psi_t(x))J_x(dy)=0,
 \qquad
 \theta(a)=e^a-a-1.
\]
Its convex conjugate is
\[
 \theta^*(b)=(b+1)\log(b+1)-b,
 \qquad b>-1.
\]
Proposition 4.2 identifies the unique minimizer of the discrete action
\[
 \int_0^1\int\theta^*(j(t,x;y)-1)\nu_t(dx)J_x(dy)dt
\]
under the forward master equation and endpoint constraints.  The proof is the same Fenchel/HJB
completion as in Proposition 4.1, using
\[
 \theta(a)+\theta^*(b)\ge ab
\]
with equality at \(b=e^a-1\).

\section{Slowing down and the optimal-transport limit}
Let \(R^k\) have generator \(L/k\), and suppose \(R^k\) obeys a path-space large-deviation
principle with speed \(\alpha_k\to\infty\) and rate \(C\).  The scaled dynamic and static problems are
\[
 \alpha_k^{-1}H(P\mid R^k),
 \qquad
 \alpha_k^{-1}H(\pi\mid R^k_{01}).
\]
Their dual objectives contain
\[
 \alpha_k^{-1}\log E_{R^k}
 \exp\{\alpha_k[\varphi(X_0)+\psi(X_1)]\}.
\]
The Laplace--Varadhan lemma gives the limit
\[
 \sup_{\omega\in\Omega}
 \{\varphi(X_0(\omega))+\psi(X_1(\omega))-C(\omega)\}.
\]
After the harmless additive normalization of \(\varphi\oplus\psi\), the limiting constraint is
\[
 \varphi(x)+\psi(y)\le c(x,y),
 \qquad
 c(x,y)=\inf\{C(\omega):\omega_0=x,\omega_1=y\}.
\]
This is the Kantorovich dual.  The survey labels the general statement informal because pointwise
dual convergence alone does not prove convergence of minimizers; the rigorous result is
Gamma-convergence.

\section{Theorem 5.2: rigorous graph limit}
For the slowed random walk, the speed is \(\log k\) and the rate counts jumps, so the induced static
cost is the graph distance \(d_\sim\).  Define
\begin{align*}
 I^k(P)&=(\log k)^{-1}H(P\mid R^k)+
  \iota_{\{P_0=\mu_0,P_1=\mu_1\}},\\
 J^k(\pi)&=(\log k)^{-1}H(\pi\mid R^k_{01})+
  \iota_{\{\pi_0=\mu_0,\pi_1=\mu_1\}}.
\end{align*}
Theorem 5.2, quoted from L\'eonard's companion work, states that these families are equicoercive and
Gamma-converge to the dynamic and static Monge--Kantorovich functionals.  Hence optimal values
converge, and every cluster point of Schr\"odinger minimizers is an optimal-transport minimizer.
The source further states uniqueness of the selected limit: among all dynamic optimal-transport
laws, the slowed bridges select the unique minimizer of entropy relative to the geodesic reference
measure obtained as the limit of \(R^{k,xy}\).

For Brownian slowing, the corresponding rate is
\(C(\omega)=\tfrac12\int_0^1|\dot\omega_t|^2dt\), the endpoint cost is
\(c(x,y)=\tfrac12|x-y|^2\), and the reference bridges concentrate on constant-speed line segments.
The survey refers the rigorous diffusion Gamma-convergence to separate work.

\section{Source-faithfulness remarks}
\begin{enumerate}[leftmargin=2em]
\item The survey distinguishes general sigma-finite entropy, static/dynamic reduction, and the
stronger product-form theorem.  These are not one theorem and are kept separate here.
\item The product form under Theorem 2.8 depends on internality and domination assumptions.  The
survey explicitly warns that without internality the density can have a more complicated support
structure.
\item The Markov property is proved by an entropy-improving conditional-independence replacement,
not merely inferred from the endpoint factorization.
\item Section 5's dual calculation is heuristic until supplemented by the cited Gamma-convergence
theorems.  The distinction is retained.
\end{enumerate}

\input{references_manual}
\end{document}
