\begin{thebibliography}{99}
\bibitem{ChenGeorgiouPavon2021} Yongxin Chen, Tryphon T. Georgiou, and Michele Pavon. \emph{Stochastic Control Liaisons: Richard Sinkhorn Meets Gaspard Monge on a Schr\"odinger Bridge}. SIAM Review, 63(2):249--313, 2021. doi:10.1137/20M1339982. arXiv:2005.10963.
\bibitem{Leonard2014Survey} Christian L\'eonard. \emph{A Survey of the Schr\"odinger Problem and Some of Its Connections with Optimal Transport}. Discrete and Continuous Dynamical Systems--A, 34(4):1533--1574, 2014. doi:10.3934/dcds.2014.34.1533. arXiv:1308.0215.
\bibitem{Leonard2019} Christian L\'eonard. \emph{Revisiting Fortet\'s Proof of Existence of a Solution to the Schr\"odinger System}. arXiv:1904.13211, 2019.
\bibitem{Fortet1940} Robert Fortet. \emph{Resolution d'un systeme d'equations de M. Schrodinger}. Journal de mathematiques pures et appliquees, 9e serie, 19(1--4):83--105, 1940.
\bibitem{SinkhornKnopp1967} Richard Sinkhorn and Paul Knopp. \emph{Concerning Nonnegative Matrices and Doubly Stochastic Matrices}. Pacific Journal of Mathematics, 21(2):343--348, 1967.
\bibitem{FranklinLorenz1989} Joel Franklin and Jens Lorenz. \emph{On the Scaling of Multidimensional Matrices}. Linear Algebra and its Applications, 114/115:717--735, 1989. doi:10.1016/0024-3795(89)90490-4.
\bibitem{Carroll2004} Joseph E. Carroll. \emph{Birkhoff's Contraction Coefficient}. Linear Algebra and its Applications, 389:227--234, 2004. doi:10.1016/j.laa.2004.02.039.
\bibitem{EvesonNussbaum1995} Simon P. Eveson and Roger D. Nussbaum. \emph{An Elementary Proof of the Birkhoff--Hopf Theorem}. Mathematical Proceedings of the Cambridge Philosophical Society, 117(1):31--55, 1995.
\bibitem{BlanchetMurthy2019} Jose Blanchet and Karthyek R. A. Murthy. \emph{Quantifying Distributional Model Risk via Optimal Transport}. Mathematics of Operations Research, 44(2):565--600, 2019. doi:10.1287/moor.2018.0936. arXiv:1604.01446.
\bibitem{GaoKleywegt2023} Rui Gao and Anton J. Kleywegt. \emph{Distributionally Robust Stochastic Optimization with Wasserstein Distance}. Mathematics of Operations Research, 48(2):603--655, 2023. doi:10.1287/moor.2022.1275. arXiv:1604.02199.
\bibitem{MohajerinEsfahaniKuhn2018} Peyman Mohajerin Esfahani and Daniel Kuhn. \emph{Data-Driven Distributionally Robust Optimization Using the Wasserstein Metric: Performance Guarantees and Tractable Reformulations}. Mathematical Programming, 171(1--2):115--166, 2018. doi:10.1007/s10107-017-1172-1. arXiv:1505.05116.
\bibitem{WangGaoXie2025} Jie Wang, Rui Gao, and Yao Xie. \emph{Sinkhorn Distributionally Robust Optimization}. Operations Research, 2025. doi:10.1287/opre.2023.0294. arXiv:2109.11926.
\bibitem{Girsanov1960} I. V. Girsanov. \emph{On Transforming a Certain Class of Stochastic Processes by Absolutely Continuous Substitution of Measures}. Theory of Probability and Its Applications, 5(3):285--301, 1960. doi:10.1137/1105027.
\bibitem{CameronMartin1945} R. H. Cameron and W. T. Martin. \emph{Transformations of Wiener Integrals under a General Class of Linear Transformations}. Transactions of the American Mathematical Society, 58(2):184--219, 1945.
\bibitem{Yeh1978} J. Yeh. \emph{Transformation of Conditional Wiener Integrals under Translation and the Cameron--Martin Translation Theorem}. Tohoku Mathematical Journal, 30(4):505--515, 1978.
\bibitem{Kakutani1948} Shizuo Kakutani. \emph{On Equivalence of Infinite Product Measures}. Annals of Mathematics, 49(1):214--224, 1948.
\bibitem{ShiDeBortoliCampbellDoucet2023} Yuyang Shi, Valentin De Bortoli, Andrew Campbell, and Arnaud Doucet. \emph{Diffusion Schr\"odinger Bridge Matching}. Advances in Neural Information Processing Systems, 2023. arXiv:2303.16852.
\bibitem{LiuLipmanNickelKarrerTheodorouChen2024} Guan-Horng Liu, Yaron Lipman, Maximilian Nickel, Brian Karrer, Evangelos A. Theodorou, and Ricky T. Q. Chen. \emph{Generalized Schr\"odinger Bridge Matching}. International Conference on Learning Representations, 2024. arXiv:2310.02233.
\bibitem{DomingoEnrichDrozdzalKarrerChen2025} Carles Domingo-Enrich, Michal Drozdzal, Brian Karrer, and Ricky T. Q. Chen. \emph{Adjoint Matching: Fine-Tuning Flow and Diffusion Generative Models with Memoryless Stochastic Optimal Control}. arXiv preprint arXiv:2409.08861, 2025.

\bibitem{Birkhoff1957} Garrett Birkhoff. \emph{Extensions of Jentzsch's Theorem}. Transactions of the American Mathematical Society, 85(1):219--227, 1957.
\bibitem{Sinkhorn1964} Richard Sinkhorn. \emph{A Relationship Between Arbitrary Positive Matrices and Doubly Stochastic Matrices}. Annals of Mathematical Statistics, 35(2):876--879, 1964.
\bibitem{Sinkhorn1967Prescribed} Richard Sinkhorn. \emph{Diagonal Equivalence to Matrices with Prescribed Row and Column Sums}. American Mathematical Monthly, 74(4):402--405, 1967.
\bibitem{Bauer1965} Friedrich L. Bauer. \emph{An Elementary Proof of the Hopf Inequality for Positive Operators}. Numerische Mathematik, 7:331--337, 1965.
\bibitem{Bushell1973} P. J. Bushell. \emph{Hilbert's Metric and Positive Contraction Mappings in a Banach Space}. Archive for Rational Mechanics and Analysis, 52:330--338, 1973.

\bibitem{CameronMartin1944Translations} R. H. Cameron and W. T. Martin. \emph{On Transformations of Wiener Integrals under Translations}. Annals of Mathematics, 45(2):386--396, 1944.
\bibitem{Luenberger1984} David G. Luenberger. \emph{Linear and Nonlinear Programming}. Second edition, Addison--Wesley, 1984.
\bibitem{RogersWilliams2000} L. C. G. Rogers and David Williams. \emph{Diffusions, Markov Processes and Martingales, Volume 2: It\^o Calculus}. Cambridge University Press, 2000.
\bibitem{PalmowskiRolski2002} Zbigniew Palmowski and Tomasz Rolski. \emph{A Technique for Exponential Change of Measure for Markov Processes}. Bernoulli, 8(6):767--785, 2002.
\bibitem{Leonard2012Girsanov} Christian L\'eonard. \emph{Girsanov Theory under a Finite Entropy Condition}. S\'eminaire de Probabilit\'es XLIV, pages 429--465, 2012.
\bibitem{LeonardRoellyZambrini2014} Christian L\'eonard, Sylvie R\"olly, and Jean-Claude Zambrini. \emph{Reciprocal Processes: A Measure-Theoretical Point of View}. Probability Surveys, 11:237--269, 2014.
\bibitem{BogachevKrasovitskiiShaposhnikov2021} Vladimir I. Bogachev, T. I. Krasovitskii, and S. V. Shaposhnikov. \emph{On Uniqueness of Probability Solutions of the Fokker--Planck--Kolmogorov Equation}. Sbornik: Mathematics, 212(6):745--781, 2021.
\bibitem{Csiszar1975} Imre Csisz\'ar. \emph{I-Divergence Geometry of Probability Distributions and Minimization Problems}. Annals of Probability, 3(1):146--158, 1975.
\bibitem{Ruschendorf1995} Ludger R\"uschendorf. \emph{Convergence of the Iterative Proportional Fitting Procedure}. Annals of Statistics, 23(4):1160--1174, 1995.
\bibitem{vanErvenHarremoes2014} Tim van Erven and Peter Harremo\"es. \emph{R\'enyi Divergence and Kullback--Leibler Divergence}. IEEE Transactions on Information Theory, 60(7):3797--3820, 2014.
\bibitem{DeBortoliThorntonHengDoucet2021} Valentin De Bortoli, James Thornton, Jeremy Heng, and Arnaud Doucet. \emph{Diffusion Schr\"odinger Bridge with Applications to Score-Based Generative Modeling}. Advances in Neural Information Processing Systems, 2021.

\bibitem{Anderson1982} Brian D. O. Anderson. \emph{Reverse-Time Diffusion Equation Models}. Stochastic Processes and their Applications, 12(3):313--326, 1982.
\bibitem{YongZhou1999} Jiongmin Yong and Xun Yu Zhou. \emph{Stochastic Controls: Hamiltonian Systems and HJB Equations}. Springer, 1999.
\bibitem{MazzoloMonthus2022} Alain Mazzolo and C\'ecile Monthus. \emph{Conditioning Diffusion Processes with Killing Rates}. Journal of Statistical Mechanics: Theory and Experiment, 2022(8):083207, 2022.
\bibitem{Kappen2005} Hilbert J. Kappen. \emph{Path Integrals and Symmetry Breaking for Optimal Control Theory}. Journal of Statistical Mechanics: Theory and Experiment, 2005(11):P11011, 2005.
\bibitem{TheodorouBuchliSchaal2010} Evangelos Theodorou, Jonas Buchli, and Stefan Schaal. \emph{A Generalized Path Integral Control Approach to Reinforcement Learning}. Journal of Machine Learning Research, 11:3137--3181, 2010.
\bibitem{Theodorou2015} Evangelos Theodorou. \emph{Nonlinear Stochastic Control and Information-Theoretic Dualities: Connections, Interdependencies and Thermodynamic Interpretations}. Entropy, 17(5):3352--3375, 2015.
\bibitem{SarkkaSolin2019} Simo S\"arkk\"a and Arno Solin. \emph{Applied Stochastic Differential Equations}. Cambridge University Press, 2019.
\bibitem{AmbrosioGigliSavare2008} Luigi Ambrosio, Nicola Gigli, and Giuseppe Savar\'e. \emph{Gradient Flows in Metric Spaces and in the Space of Probability Measures}. Second edition, Birkh\"auser, 2008.
\bibitem{LipmanEtAl2023} Yaron Lipman, Ricky T. Q. Chen, Heli Ben-Hamu, Maximilian Nickel, and Matt Le. \emph{Flow Matching for Generative Modeling}. International Conference on Learning Representations, 2023.
\bibitem{LiuEtAl2022DeepGSB} Guan-Horng Liu, Tianrong Chen, Oswin So, and Evangelos A. Theodorou. \emph{Deep Generalized Schr\"odinger Bridge}. Advances in Neural Information Processing Systems, 2022.
\bibitem{ChenGeorgiou2015LinearBridges} Yongxin Chen and Tryphon T. Georgiou. \emph{Stochastic Bridges of Linear Systems}. IEEE Transactions on Automatic Control, 61(2):526--531, 2016.
\bibitem{KappenRuiz2016} Hilbert J. Kappen and Hans Christian Ruiz. \emph{Adaptive Importance Sampling for Control and Inference}. Journal of Statistical Physics, 162(5):1244--1266, 2016.
\bibitem{NeklyudovEtAl2023} Kirill Neklyudov, Daniel Severo, and Alireza Makhzani. \emph{Action Matching: A Variational Method for Learning Stochastic Dynamics from Samples}. International Conference on Machine Learning, 2023.
\bibitem{PaleyWiener1934} Raymond E. A. C. Paley and Norbert Wiener. \emph{Fourier Transforms in the Complex Domain}. American Mathematical Society Colloquium Publications, volume 19, 1934.
\bibitem{Sunouchi1951} G. Sunouchi. \emph{Harmonic Analysis and Wiener Integrals}. Tohoku Mathematical Journal, 3:187--196, 1951.
\bibitem{Segal1958} Irving E. Segal. \emph{Distributions in Hilbert Space and Canonical Systems of Operators}. Transactions of the American Mathematical Society, 88:12--41, 1958.
\end{thebibliography}
