\IfFileExists{papers/formalized-vs-literature-preamble.tex}{\input{papers/formalized-vs-literature-preamble}}{\input{formalized-vs-literature-preamble}}
\title{Blanchet--Murthy Theorem 1: formalized result versus the literature}
\author{}
\date{Formalization snapshot \snapshot}
\begin{document}
\maketitle
\statusline{\Class{A}: weak-duality proof plus an explicit reverse-inequality premise}{The difficult minimax/reverse-duality direction, dual attainment, and complementary slackness are not derived from Assumptions (A1)--(A2).}
\section{Source theorem}
Blanchet and Murthy, ``Quantifying Distributional Model Risk via Optimal Transport,'' Theorem~1(a) proves equality between the robust primal and an infinite-dimensional dual; Theorem~1(b) gives a dual optimizer and complementary-slackness relations.  Remark~1 reduces the value to the scalar dual
\[
 \inf_{\lambda\ge0}\left\{\lambda\delta+
 \int \sup_y(f(y)-\lambda c(x,y))\,\widehat\mu(dx)\right\}.
\]
The complete local reconstruction is
\RepoFile{prose/distilled_literature/BlanchetMurthy2019_theorem1.tex}.

\section{Lean declarations}
\Lean{BlanchetMurthy2019.wdro_strong_duality} and
\Lean{BlanchetMurthy2019.wdro_strong_duality_dualFn} are in
\RepoFile{BlanchetMurthy2019/Basic.lean}.

\section{What Lean proves internally}
The proof derives the full weak-duality direction from the reusable per-coupling Lagrangian inequality, handles the source/scaffold orientation through an explicit symmetry assumption, derives boundedness of the primal expectation set, and performs the real $sSup/sInf$ bookkeeping.

The theorem then accepts
\[
 \text{dual value}\le \text{primal value}
\]
as an explicit hypothesis \texttt{hge}.  Consequently the final equality is machine checked, but the source theorem has not been derived from its primitive topological hypotheses.

\section{Axis-by-axis comparison}
\begin{comparison}
Cost assumptions & Source: nonnegative lower-semicontinuous cost with identity-of-indiscernibles. Lean: arbitrary real cost in the type, plus symmetry and explicit finiteness/integrability premises.\\
Reward assumptions & Source: upper semicontinuous and $L^1$ under the nominal law. Lean: explicit envelope boundedness and integrability under all feasible laws.\\
Reverse inequality & Source: proved by the paper's minimax argument. Lean: supplied as \texttt{hge}.\\
Dual attainment & Source: yes. Lean: absent.\\
Complementary slackness & Source: yes. Lean: absent.\\
Primal attainment & Source: conditional and discussed separately. Lean: absent.\\
\end{comparison}

\section{Relationship to the Gao--Kleywegt route}
The repository contains a genuinely proved reverse-duality theorem under stronger bounded-continuous assumptions in \RepoFile{ForMathlib/OptimalTransport/StrongDualityGe.lean}.  A future Blanchet--Murthy completion may reuse that infrastructure after generalizing it, or it may formalize the source's Fenchel--Sion proof to obtain the stronger optimizer and complementary-slackness conclusions.
\end{document}
