\IfFileExists{papers/formalized-vs-literature-preamble.tex}{\input{papers/formalized-vs-literature-preamble}}{\input{formalized-vs-literature-preamble}}
\title{Kolmogorov extension and Wiener reference measure: formalized result versus the literature}
\author{}
\date{Formalization snapshot \snapshot}
\begin{document}
\maketitle
\statusline{\Class{G/V}: concrete Wiener measure on the unrestricted coordinate path space built from a vendored extension theorem}{The measure lives on all functions $\R_{\ge0}\to\R$; concentration on a canonical continuous anchored path space and its Borel generation are not yet formalized.}
\section{Result}
\Lean{ForMathlib.MeasureTheory.wienerMeasure} in
\RepoFile{ForMathlib/MeasureTheory/WienerMeasure.lean} applies a Kolmogorov extension theorem to Mathlib's consistent Brownian Gaussian projective family.  The repository proves that it is a probability measure and that every finite coordinate restriction has the expected centered Gaussian law with covariance $\min(s,t)$.

\section{Provenance}
The projective-limit implementation under \RepoFile{ForMathlib/KolmogorovExtension/} is vendored from Remy Degenne's
\url{https://github.com/RemyDegenne/kolmogorov_extension4} under Apache-2.0; exact commit metadata is recorded in the vendored headers/README.  The assembly using Mathlib's Brownian projective family is local.

\section{Literature comparison}
Kolmogorov's extension theorem produces a measure on a coordinate product.  The classical Wiener construction then proves continuity of sample paths (or constructs directly on a continuous-path carrier after modification).  The current Lean result completes the projective-limit step but not the continuity/support step.

\begin{comparison}
Index set & Nonnegative real times.\\
Carrier & Unrestricted functions $\R_{\ge0}\to\R$.\\
Finite-dimensional laws & Fully proved.\\
Probability mass & Fully proved.\\
Continuous-path concentration & Open.\\
Anchored path type & Open/provisional.\\
Dyadic Borel generation & Open.\\
\end{comparison}

\section{Role in DRSB}
This gives a concrete reference law for finite-dimensional projection arguments.  It does not by itself support path-level Cameron--Martin translation, disintegration into Brownian bridges, SDE laws, or Girsanov.
\end{document}
